---
title: "Final-full-mouse-analysis-combat"
author: "Elisa Guma"
date: "3/15/2023"
output: html_document
---
### Description

Here I will analyze the full sample of mouse controls to get a first pass at sex differences in the mouse brain.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 300)
library(RMINC)
library(data.tree)
library(grid)
library(PupillometryR)
library(MRIcrotome)
library(tidyverse)
library(reshape2)
library(car)
library(lme4)
library(lmerTest)
source('../input_data/tree_tools.R')
library(sva)
```


# Description
In this script I aim to compute maps of sex differences in the brain using in a 
large collection of WT mice on a C57BL6/J or N background.

I characterize differences sex differences in total and regional brain volume (both
with and without total brain volume covariation). In addition to mean differences, I 
also investigate sex differences in anatomical variance. 

# Import demographics data
The data exists in two forms - an RData file containing the volumes, and a CSV file describing each scan. Let's load both:
```{r}
volumes <- load("../input_data/Volumes_150.RData")
gf <- read_csv("../input_data/mouse_demographics.csv")
```
Check that you tree volume dimensions match the demographics spreadsheet
```{r}
dim(gf)
dim(vols)
```

# Set up hierarchical atlas
```{r make-the-tree}
# this file contains the definitions of all our labels
defs <- "../mouse_anatomy/input_data/DSURQE_40micron_R_mapping.csv"

# this file contains the Allen institutes hierarchical definitions
abijson <- "../mouse_anatomy/Allen_hierarchy_definitions.json"

# now we combine our label definitions with the hierarchical structure
hdefs_tree <- makeMICeDefsHierachical(defs, abijson) 

# and then we add the volume information to those hierarchies
hdefs_tree <- addVolumesToHierarchy(hdefs_tree, vols)
```

## Read in the label and anatomical volumes
```{r} 
labelVol <- mincArray(mincGetVolume("../mouse_anatomy/DSURQE_40micron_labels.mnc"))
anatVol <- mincArray(mincGetVolume("../mouse_anatomy/DSURQE_40micron_average.mnc"))
```

## Add volumes to hierarchical tree
With the hierarchies established, it’s time to add the volumes to hierarchical tree.
```{r}
hdefs_tree <- addVolumesToHierarchy(hdefs_tree, vols)
class(hdefs_tree)
labelCols <- hanatToVolume(hdefs_tree, labelVol, "color_hex_triplet") #load in the Allen atlas colours to display next to the results
```

## Extract volumes into data fram
```{r make-final-data}
df_volumes <- hdefs_tree$Get("volumes") %>% as_tibble
gf$ventricle_volume <- df_volumes$`ventricular systems`
gf$TBV <- df_volumes$`root2`
gf$total_tissue_volume <- gf$TBV - gf$ventricle_volume
```

## Prune white matter and ventricles out of tree
### Whole tree
```{r}
#Clone the tree so that any changes you make do not affect your original tree
hdefsCopy <- Clone(hdefs_tree)
nodes_to_cut <- c("ventricular systems", "fiber tracts") #remove ventricles and cranial nerves
pruneAnatTree(hdefsCopy, nodes = nodes_to_cut, method = "AtNode")
df_volumes_pruned <- hdefsCopy$Get("volumes") %>% as_tibble
```
### Extract just leaves from data 
```{r}
df_volumes_pruned_leaves <- hdefsCopy$Get("volumes", filterFun = isLeaf) %>% as_tibble
```

### Assign colour palette for positive and negative beta values 
```{r}
positive_cols <- function() {
  getOption("MRIcrotomeCol", colorRampPalette(c("yellow","red"))(255))
}
negative_cols <- function() {
  getOption("MRIcrotomeRcol", colorRampPalette(c("turquoise3","royalblue"))(255))
}
labelCols <- hanatToVolume(hdefsCopy, labelVol, "color_hex_triplet") #load in the Allen atlas colours to display next to the results
```

## Filter to WT only 
### Whole tree
```{r}
pre_export <- gf %>% dplyr::select(Mouse_ID, Is_Wildtype, Background,
                        Study_Name, Mouse_Sex, Mouse_Age, Timepoint, Treatment_Code, TBV, total_tissue_volume) %>%
          bind_cols(df_volumes_pruned) %>% distinct(Mouse_ID, .keep_all = TRUE) 

export1 <- pre_export %>% 
          dplyr::mutate(Mouse_Age = str_remove(Mouse_Age,"P")) %>%
          separate(Mouse_Age, c("a", "b"), fill = "right") %>%
          dplyr::mutate(a = as.numeric(a), b = as.numeric(b)) %>%
          dplyr::mutate(Mouse_Age = ifelse(is.na(b) == FALSE, (a + b)/2, 
                                    a), .after = Mouse_Sex) %>% dplyr::select(-c(a,b))

#Replace NAs with age of 60 (standard age for the mice acquired in the studies)
export1["Mouse_Age"][is.na(export1["Mouse_Age"])] <- 60 
export <- export1 %>% filter(Mouse_Age < 200)
#print(export$Mouse_Age)
export$AGE_cent <- export$Mouse_Age - mean(export$Mouse_Age)
```

## Subset data such that there are at least 5M and 5F per group
### Whole tree
```{r}
nLimit <- 4
balanced_groups_by_cohort5_1 <-  export %>% filter(Is_Wildtype == "WT" & !duplicated(Mouse_ID) & Treatment_Code %in% c("none", "MOCK", "VEH", "Control","Saline") &
                                                     Background %in% c("C57BL-6J", "C57BL-6N", "C57BL6-N","C57BL-6"))
balanced_groups_by_cohort5 <- balanced_groups_by_cohort5_1 %>%  
  dplyr::group_by(Study_Name) %>% 
  dplyr::summarize(nF=sum(Mouse_Sex == "F"),
            nM=sum(Mouse_Sex == "M"),
             n = n()) %>%
  dplyr::filter(nF > nLimit & nM > nLimit)
balanced_groups_by_cohort5

balanced_groups_by_cohort5 <- balanced_groups_by_cohort5_1 %>% 
  filter (Study_Name %in% balanced_groups_by_cohort5$Study_Name)

table(balanced_groups_by_cohort5$Background, balanced_groups_by_cohort5$Study_Name)
table(balanced_groups_by_cohort5$Background, balanced_groups_by_cohort5$Mouse_Sex)
table(balanced_groups_by_cohort5$Study_Name, balanced_groups_by_cohort5$Mouse_Sex)
table(balanced_groups_by_cohort5$Background,balanced_groups_by_cohort5$Study_Name, balanced_groups_by_cohort5$Mouse_Sex)

# grab the distinct cohort names from the cohorts data 
cohort_names <- unique(balanced_groups_by_cohort5$Study_Name)

# look at mean age per sex and sample size
demographics_table5 <- balanced_groups_by_cohort5 %>% 
  dplyr::group_by(Mouse_Sex) %>% 
  dplyr::summarize(
    mean_age = mean(Mouse_Age),
    sd_age = sd(Mouse_Age),
    range_age=range(Mouse_Age),
    n = n())
demographics_table5

# export csv
balanced_groups_by_cohort5$Background <- gsub("C57BL6-N", "C57BL-6N", balanced_groups_by_cohort5$Background)
write.csv(balanced_groups_by_cohort5, file="/Users/gumae2/Documents/Cross_species_sex_differences/mouse_data/input_data/WT_mouse_data_C57.csv")
```

## Subset the data so it's just C57BL6J and 6N
```{r}
df_just_C57BL6J <-balanced_groups_by_cohort5 %>% filter(Background == "C57BL-6J")
table(df_just_C57BL6J$Study_Name, df_just_C57BL6J$Mouse_Sex)
chisq.test(df_just_C57BL6J$Mouse_Sex, df_just_C57BL6J$Study_Name)

df_just_C57BL6N <-balanced_groups_by_cohort5 %>% filter(Background %in% c("C57BL-6N","C57BL6-N"))
table(df_just_C57BL6N$Study_Name, df_just_C57BL6N$Mouse_Sex)
chisq.test(df_just_C57BL6N$Mouse_Sex, df_just_C57BL6N$Study_Name)
```


# Apply ComBat harmonization
Here we separate the background strains and apply ComBat harmonization within each, separately,
to account for any potential study specific variance that we are not insterested in modeling.
```{r}
# C57BL6J
mouse_vols_C57BL6J_t <- t(as.matrix(df_just_C57BL6J[,10:464]))
mouse_demo_C57BL6J <- df_just_C57BL6J[,c(1:9,465)]

cohort_batch_C57BL6J <- df_just_C57BL6J$Study_Name
combat_vols_cohort_C57BL6J <- sva::ComBat(dat = mouse_vols_C57BL6J_t, batch = cohort_batch_C57BL6J)

combat_vols_C57BL6J_reoriented <- t(combat_vols_cohort_C57BL6J)
df_combat_C57BL6J_vols <- mouse_demo_C57BL6J %>%  cbind(combat_vols_C57BL6J_reoriented) %>% as.data.frame()

# C57BL6N
mouse_vols_C57BL6N_t <- t(as.matrix(df_just_C57BL6N[,10:464]))
mouse_demo_C57BL6N <- df_just_C57BL6N[,c(1:9,465)]

cohort_batch_C57BL6N <- df_just_C57BL6N$Study_Name
combat_vols_cohort_C57BL6N <- sva::ComBat(dat = mouse_vols_C57BL6N_t, batch = cohort_batch_C57BL6N)

combat_vols_C57BL6N_reoriented <- t(combat_vols_cohort_C57BL6N)
df_combat_C57BL6N_vols <- mouse_demo_C57BL6N %>%  cbind(combat_vols_C57BL6N_reoriented) %>% as.data.frame()

df_combat_vols <- rbind(df_combat_C57BL6J_vols, df_combat_C57BL6N_vols)
```
## Demographics Analysis
```{r}
# look at mean age per sex and sample size
demographics_cobmat <- df_combat_vols %>% 
  dplyr::group_by(Mouse_Sex) %>% 
  dplyr::summarize(
    mean_age = mean(Mouse_Age),
    sd_age = sd(Mouse_Age),
    range_age=range(Mouse_Age),
    n = n())
demographics_cobmat

summary(aov(Mouse_Age ~ Mouse_Sex, df_combat_vols))
chisq.test(df_combat_vols$Mouse_Sex, df_combat_vols$Study_Name)
```

## Calculate percent difference in volume for total tissue volume
```{r}
mean_difference <- df_combat_vols %>% 
  dplyr::group_by(Mouse_Sex) %>% 
  dplyr::summarize(
    mean_ttv = mean(total_tissue_volume),
    sd_ttv=sd(total_tissue_volume),
    n = n()
  )
mean_difference

female_mean <-430.4453		
male_mean <- 431.8090
pct_diff = ((male_mean - female_mean)/((male_mean + female_mean)/2))*100
pct_diff
```

### Z-score volumes following ComBat harmonization
```{r}
mouse_vols_combat <- df_combat_vols[,11:ncol(df_combat_vols)] %>% as.matrix()
mouse_demo_combat <- df_combat_vols[,c(1:10)]

zscores_volumes_combat <- matrix(0, nrow = nrow(mouse_vols_combat), ncol = ncol(mouse_vols_combat))
for(j in 1:ncol(mouse_vols_combat)){
  mu <- mean(mouse_vols_combat[,j]) #Then you can compute the ROI-wise averages
  sigma <- sd(mouse_vols_combat[,j]) #And standard deviations
  zscores_volumes_combat[,j] <- (mouse_vols_combat[,j] - mu)/sigma #Then you can use those to make the z-score volumes
}
colnames(zscores_volumes_combat)<-colnames(mouse_vols_combat)
df_zscored_data_combat <- mouse_demo_combat %>%  cbind(zscores_volumes_combat) %>% as.data.frame()


### Check for outliers
df_data_no_outliers_combat <- df_zscored_data_combat %>% dplyr::mutate(across(!c("Mouse_ID","Is_Wildtype","Background",
                                                                        "Study_Name","Mouse_Sex","Mouse_Age","Timepoint", "AGE_cent",
                                                                        "Treatment_Code"), ~ ifelse(.x <= -4*sd(.x) + mean(.x), NA, 
                                             ifelse(.x >= 4*sd(.x) + mean(.x), NA, .x)))) 
```

No outliers were detected so we proceed.

# FULL SAMPLE (ComBat harmonized)

## Total tissue volume
### Linear Model
```{r}
linearmodel <- lm(total_tissue_volume ~ Mouse_Sex+Background + AGE_cent, data=df_zscored_data_combat)
summary(linearmodel)

summary(df_zscored_data_combat$total_tissue_volume)
```

## Variance in total tissue volume
```{r}
# Run leveneTest() to check for differences in variance
result = leveneTest(total_tissue_volume ~ Mouse_Sex, df_zscored_data_combat)
print(result)

ggplot(df_zscored_data_combat, aes(x=total_tissue_volume, fill=Mouse_Sex, color=Mouse_Sex)) +
    geom_density(alpha=.4, adjust = 0.5) +
    theme_classic()+ xlab("Total Tissue Volume")+ylab("Density")+
    scale_fill_manual(values=c("F"="deepskyblue","M"="goldenrod1"))+
    scale_colour_manual(values=c("F"="deepskyblue","M"="goldenrod1"))+
   theme(text = element_text(size = 14))+
    geom_vline(xintercept=c(0))

summary_table <- df_zscored_data_combat %>% 
  group_by(Mouse_Sex) %>%
  dplyr::summarize(
    mean_ttv = mean(`total_tissue_volume`),
    median_ttv = median(`total_tissue_volume`),
    sd_ttv=sd(`total_tissue_volume`),
    range_ttv=range(`total_tissue_volume`),
    n = n()
  )
summary_table
```

### Plot total tissue volume
```{r}
## zscore total tissue volume
mu_ttv<- mean(df_combat_vols$total_tissue_volume)
sigma_ttv <- sd(df_combat_vols$total_tissue_volume)

ggplot(df_zscored_data_combat, aes(x = Mouse_Sex, y = total_tissue_volume, group=Mouse_Sex, colour=Mouse_Sex, fill=Mouse_Sex)) + 
geom_flat_violin(position = position_nudge(x = .1, y = 0), alpha =.4,adjust =0.8)+
  geom_boxplot(width = .1, outlier.shape = NA,alpha = .5) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(side = "l",range_scale = .4,alpha = .3) +
  #Adjust theme
  scale_fill_manual(values=c("deepskyblue", "goldenrod1"))+ 
  scale_color_manual(values = c("deepskyblue", "goldenrod1"))+
  theme_classic()+
  labs(
    y="Total Tissue Volume (Z-scored)",
    fill="Mouse_Sex")+  theme(text = element_text(size = 14))+ylim(-4, 4)+
  scale_y_continuous(
    sec.axis = sec_axis(name=expression(Total ~ Tissue ~ Volume~ (mm^3)), ~ (. * sigma_ttv) + mu_ttv), 
    name="Total Tissue Volume (Z-scored)", limits=c(-4,4)) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off")

pdf(file = "/Users/gumae2/Documents/Cross_species_sex_differences/Figures/total_tissue_volume/mouse_ttv_split_by_grp_c57_ComBat.pdf",   # The directory you want to save the file in
    width = 8.5, # The width of the plot in inches
    height = 4.5) # The height of the plot in inches
ggplot(df_data_no_outliers_combat, aes(x = Mouse_Sex , y = total_tissue_volume, group=Mouse_Sex, colour=Study_Name)) + 
  geom_boxplot(width = .5, outlier.shape = NA,alpha = .5) +
  geom_point() +geom_jitter(width = 0.25)+
  facet_grid(~Background) +
  ## add justified jitter from the {gghalves} package
  theme_classic()+
  labs(
    y="Total Tissue Volume (Z-scored)",
    fill="Mouse_Sex")+  theme(text = element_text(size = 14))+
  scale_y_continuous(
    sec.axis = sec_axis(name=expression(Total ~ Tissue ~ Volume~ (mm^3)), ~ (. * sigma_ttv) + mu_ttv), 
    name="Total Tissue Volume (Z-scored)", limits=c(-4,4)) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off")

# Look at Background Strain separately
ggplot(df_zscored_data_combat, aes(x = Mouse_Sex , y = total_tissue_volume, group=Mouse_Sex, colour=Mouse_Sex)) + 
  geom_boxplot(width = .5, outlier.shape = NA,alpha = .5) +
  geom_point() +geom_jitter(width = 0.25)+
  facet_grid(~Background) +
  ## add justified jitter from the {gghalves} package
  theme_classic()+
  labs(
    y="Total Tissue Volume (Z-scored)",
    fill="Mouse_Sex")+  theme(text = element_text(size = 14))+ylim(-4, 4)+
  scale_fill_manual(values=c("deepskyblue", "goldenrod1"))+ 
  scale_color_manual(values = c("deepskyblue", "goldenrod1"))+
  scale_y_continuous(
    sec.axis = sec_axis(name=expression(Total ~ Tissue ~ Volume~ (mm^3)), ~ (. * sigma_ttv) + mu_ttv), 
    name="Total Tissue Volume (Z-scored)", limits=c(-4,4)) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off")
```

## Prune white matter and ventricles out of tree to look at sex differences
```{r}
#Clone the tree so that any changes you make do not affect your original tree
hdefs_GM_WM <- Clone(hdefs_tree)
df_volumes_GMWM <- hdefs_GM_WM$Get("volumes") %>% as_tibble

pre_export_GMWM <- gf %>% dplyr::select(Mouse_ID, Is_Wildtype, POND_Mouse_ID, Background,
                        Study_Name,Mouse_Sex, Mouse_Age, Timepoint, Treatment_Code, TBV, total_tissue_volume) %>%
          bind_cols(df_volumes_GMWM) %>% distinct(Mouse_ID, .keep_all = TRUE) 

export_GMWM1 <- pre_export_GMWM %>% 
          mutate(Mouse_Age = str_remove(Mouse_Age,"P")) %>%
          separate(Mouse_Age, c("a", "b"), fill = "right") %>%
          mutate(a = as.numeric(a), b = as.numeric(b)) %>%
          mutate(Mouse_Age = ifelse(is.na(b) == FALSE, (a + b)/2, 
                                    a), .after = Mouse_Sex) %>%
          dplyr::select(-c(a,b)) 
#Replace NAs with age of 60
export_GMWM1["Mouse_Age"][is.na(export_GMWM1["Mouse_Age"])] <- 60
export_GMWM <- export_GMWM1 %>% filter(Mouse_Age < 200)
export_GMWM$AGE_cent <- export_GMWM$Mouse_Age - mean(export_GMWM$Mouse_Age)

nLimit <- 4
df_GMWM_1 <-  export_GMWM %>% filter(Is_Wildtype == "WT" & !duplicated(Mouse_ID) & Treatment_Code %in% c("none", "MOCK", "VEH", "Control", "Saline"))
df_GMWM <- df_GMWM_1 %>%  
  dplyr::group_by(Study_Name) %>% 
  dplyr::summarize(nF=sum(Mouse_Sex == "F"),
            nM=sum(Mouse_Sex == "M"),
             n = n()) %>%
  dplyr::filter(nF > nLimit & nM > nLimit)
df_GMWM

df_GMWM <- df_GMWM_1 %>%filter (Study_Name %in% df_GMWM$Study_Name)

# grab the distinct cohort names from the cohorts data 
cohort_names <- unique(df_GMWM$Study_Name)

## Subset the data so it's just C57BL6J and 6N
df_just_C57BL6J_forGMWM <-df_GMWM %>% filter(Background == "C57BL-6J")
df_just_C57BL6N_forGMWM <-df_GMWM %>% filter(Background %in% c("C57BL-6N", "C57BL6-N"))
df_just_C57BL6N_forGMWM$Background <- gsub("C57BL6-N", "C57BL-6N", df_just_C57BL6N_forGMWM$Background)

# Apply ComBat to each background strain
# C57BL6J
mouse_GMWM_C57BL6J_t <- t(as.matrix(df_just_C57BL6J_forGMWM[,10:601]))
mouse_GMWM_demo_C57BL6J <- df_just_C57BL6J_forGMWM[,c(1:9,602)]

cohort_batch_C57BL6J_forGMWM <- df_just_C57BL6J_forGMWM$Study_Name
combat_vols_cohort_C57BL6J_forGMWM <- sva::ComBat(dat = mouse_GMWM_C57BL6J_t, batch = cohort_batch_C57BL6J_forGMWM)

combat_vols_C57BL6J_GMWM_reoriented <- t(combat_vols_cohort_C57BL6J_forGMWM)
df_combat_C57BL6J_vols_GMWM <- mouse_GMWM_demo_C57BL6J %>%  cbind(combat_vols_C57BL6J_GMWM_reoriented) %>% as.data.frame()

# C57BL6N
mouse_GMWM_C57BL6N_t <- t(as.matrix(df_just_C57BL6N_forGMWM[,10:601]))
mouse_GMWM_demo_C57BL6N <- df_just_C57BL6N_forGMWM[,c(1:9,602)]

cohort_batch_C57BL6N_forGMWM <- df_just_C57BL6N_forGMWM$Study_Name
combat_vols_cohort_C57BL6N_forGMWM <- sva::ComBat(dat = mouse_GMWM_C57BL6N_t, batch = cohort_batch_C57BL6N_forGMWM)

combat_vols_C57BL6N_GMWM_reoriented <- t(combat_vols_cohort_C57BL6N_forGMWM)
df_combat_C57BL6N_vols_GMWM <- mouse_GMWM_demo_C57BL6N %>%  cbind(combat_vols_C57BL6N_GMWM_reoriented) %>% as.data.frame()
df_combat_vols_GMWM <- rbind(df_combat_C57BL6J_vols_GMWM, df_combat_C57BL6N_vols_GMWM)

mouse_vols_GMWM_combat <- df_combat_vols_GMWM[,12:ncol(df_combat_vols_GMWM)] %>% as.matrix()
mouse_demo_GMWM_combat <- df_combat_vols_GMWM[,c(1:11)]

zscores_volumes_combat_GMWM <- matrix(0, nrow = nrow(mouse_vols_GMWM_combat), ncol = ncol(mouse_vols_GMWM_combat))
for(j in 1:ncol(mouse_vols_GMWM_combat)){
  mu <- mean(mouse_vols_GMWM_combat[,j]) #Then you can compute the ROI-wise averages
  sigma <- sd(mouse_vols_GMWM_combat[,j]) #And standard deviations
  zscores_volumes_combat_GMWM[,j] <- (mouse_vols_GMWM_combat[,j] - mu)/sigma #Then you can use those to make the z-score volumes
}
colnames(zscores_volumes_combat_GMWM)<-colnames(mouse_vols_GMWM_combat)
df_zscores_volumes_combat_GMWM <- mouse_demo_GMWM_combat %>%  cbind(zscores_volumes_combat_GMWM) %>% as.data.frame()
```

## Calculate model outputs to plot adjusted volumes
```{r}
model_inputs_mouse <- df_zscores_volumes_combat_GMWM %>% 
  summarize(
    AGE_cent = mean(AGE_cent),
    total_tissue_volume = mean(total_tissue_volume))
model_inputs_mouse %>% as.data.frame()
model_inputs_mouse[2,] <- model_inputs_mouse[1,]
model_inputs_mouse$Mouse_Sex = c("F", "M")
```

## Plot White matter Raw
```{r}
mu_wm <- mean(df_combat_vols_GMWM$`fiber tracts`)
sd_wm <- sd(df_combat_vols_GMWM$`fiber tracts`)

ggplot(df_zscores_volumes_combat_GMWM, aes(x = Mouse_Sex, y = `fiber tracts`, group=Mouse_Sex, colour=Mouse_Sex, fill=Mouse_Sex)) + 
geom_flat_violin(
    position = position_nudge(x = .1, y = 0), 
    alpha =.4,adjust =0.8)+
  geom_boxplot(width = .1, outlier.shape = NA,alpha = .5) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(side = "l", range_scale = .4, alpha = .3) +
  #Adjust theme
  scale_fill_manual(values=c("deepskyblue", "goldenrod1"))+ 
  scale_color_manual(values = c("deepskyblue", "goldenrod1"))+
  theme_classic()+
  labs(y="White Matter Volume (Z-scored)",
    fill="Mouse_Sex")+  theme(text = element_text(size = 14))+ylim(-4, 4)+
  scale_y_continuous(
    sec.axis = sec_axis(name=expression(White ~ Matter ~ Volume~ (mm^3)), ~ (. * sd_wm) + mu_wm), 
    name="White Matter Volume (Z-scored)", limits=c(-4,4)) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off")
```
## White Matter Adjusted
```{r}
linearmodel <- lm(`fiber tracts` ~ Mouse_Sex+ total_tissue_volume + AGE_cent, data=df_zscores_volumes_combat_GMWM)
summary(linearmodel)

## estimate the residulized ROI volume for females (1) and males (2)
wm_estimates <- predict.lm(linearmodel,model_inputs_mouse)

residualized_wm_vol <- residuals(lm(`fiber tracts` ~ total_tissue_volume + AGE_cent + Background, data=df_zscores_volumes_combat_GMWM))
resid_wm<- cbind(df_zscores_volumes_combat_GMWM,residualized_wm_vol)
summary(lm(residualized_wm_vol ~ Mouse_Sex , data=resid_wm))

resid_wm_females  <- resid_wm %>% filter(Mouse_Sex %in% c("F"))
resid_wm_females$adjusted_wm_resid <- resid_wm_females$residualized_wm_vol + wm_estimates[1]

resid_wm_males  <- resid_wm %>% filter(Mouse_Sex %in% c("M"))
resid_wm_males$adjusted_wm_resid <- resid_wm_males$residualized_wm_vol + wm_estimates[2]

resid_wm_adjusted <- rbind(resid_wm_females, resid_wm_males)
```
### Plot White matter adjusted
```{r}
ggplot(resid_wm_adjusted, aes(x = Mouse_Sex, y = adjusted_wm_resid, group=Mouse_Sex, colour=Mouse_Sex, fill=Mouse_Sex)) + 
geom_flat_violin(
    position = position_nudge(x = .1, y = 0), 
    alpha =.4,adjust =0.8)+
  geom_boxplot(width = .1, outlier.shape = NA,
     alpha = .5) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(side = "l", range_scale = .4, alpha = .3) +
  #Adjust theme
  scale_fill_manual(values=c("deepskyblue", "goldenrod1"))+ 
  scale_color_manual(values = c("deepskyblue", "goldenrod1"))+
  theme_classic()+
  labs(y="White Matter Volume (residualized)",
    fill="Mouse_Sex")+  theme(text = element_text(size = 14))+
  scale_y_continuous(
    sec.axis = sec_axis(name=expression(White ~ Matter ~ Volume~ (mm^3)), ~ (. * sd_wm) + mu_wm), 
    name="White Matter Volume (residualized)") +
   coord_cartesian(xlim = c(1.2, NA), clip = "off")
```

## Plot Total Gray Matter Raw
```{r}
mu_gm <- mean(df_combat_vols_GMWM$`Basic cell groups and regions`)
sd_gm <- sd(df_combat_vols_GMWM$`Basic cell groups and regions`)

ggplot(df_zscores_volumes_combat_GMWM, aes(x = Mouse_Sex, y = `Basic cell groups and regions`, group=Mouse_Sex, colour=Mouse_Sex, fill=Mouse_Sex)) + 
geom_flat_violin(
    position = position_nudge(x = .1, y = 0), 
    alpha =.4,adjust =0.8)+
  geom_boxplot(width = .1, outlier.shape = NA,alpha = .5) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(side = "l", range_scale = .4, alpha = .3) +
  #Adjust theme
  scale_fill_manual(values=c("deepskyblue", "goldenrod1"))+ 
  scale_color_manual(values = c("deepskyblue", "goldenrod1"))+
  theme_classic()+
  labs(
    y="Grey Matter Volume (Z-scored)",
    fill="Mouse_Sex")+  theme(text = element_text(size = 14))+ylim(-4, 4)+
  scale_y_continuous(
    sec.axis = sec_axis(name=expression(Grey ~ Matter ~ Volume~ (mm^3)), ~ (. * sd_wm) + mu_wm), 
    name="Grey Matter Volume (Z-scored)", limits=c(-4,4)) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off")
```

## Grey Matter Adjusted
```{r}
linearmodel <- lm(`Basic cell groups and regions` ~ Mouse_Sex+ total_tissue_volume + AGE_cent, data=df_zscores_volumes_combat_GMWM)
summary(linearmodel)

## estimate the residulized ROI volume for females (1) and males (2)
gm_estimates <- predict.lm(linearmodel,model_inputs_mouse)

residualized_gm_vol <- residuals(lm(`Basic cell groups and regions` ~ total_tissue_volume + AGE_cent, data=df_zscores_volumes_combat_GMWM))
resid_gm<- cbind(df_zscores_volumes_combat_GMWM,residualized_gm_vol)
summary(lm(residualized_gm_vol ~ Mouse_Sex , data=resid_gm))

resid_gm_females  <- resid_gm %>% filter(Mouse_Sex %in% c("F"))
resid_gm_females$adjusted_gm_resid <- resid_gm_females$residualized_gm_vol + gm_estimates[1]

resid_gm_males  <- resid_gm %>% filter(Mouse_Sex %in% c("M"))
resid_gm_males$adjusted_gm_resid <- resid_gm_males$residualized_gm_vol + gm_estimates[2]

resid_gm_adjusted <- rbind(resid_gm_females, resid_gm_males)
```
### Plot grey matter adjusted
```{r}
ggplot(resid_gm_adjusted, aes(x = Mouse_Sex, y = adjusted_gm_resid, group=Mouse_Sex, colour=Mouse_Sex, fill=Mouse_Sex)) + 
geom_flat_violin(
    position = position_nudge(x = .1, y = 0), 
    alpha =.4,adjust =0.8)+
  geom_boxplot(width = .1, outlier.shape = NA,
     alpha = .5) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(side = "l", range_scale = .4, alpha = .3) +
  #Adjust theme
  scale_fill_manual(values=c("deepskyblue", "goldenrod1"))+ 
  scale_color_manual(values = c("deepskyblue", "goldenrod1"))+
  theme_classic()+
  labs(y="Grey Matter Volume (residualized)",
    fill="Mouse_Sex")+  theme(text = element_text(size = 14))+
  scale_y_continuous(
    sec.axis = sec_axis(name=expression(Grey ~ Matter ~ Volume~ (mm^3)), ~ (. * sd_gm) + mu_gm), 
    name="Grey Matter Volume (residualized)") +
   coord_cartesian(xlim = c(1.2, NA), clip = "off")
```

# Run a linear model across all regions (ComBat harmonized volumes)
```{r}
sexdiff_full_lm_combat <- anatLm(~ Mouse_Sex+Background+ AGE_cent + total_tissue_volume, df_zscored_data_combat, df_zscored_data_combat[,13:ncol(df_zscored_data_combat)])
FDR_sexdiff_full_lm_combat <- anatFDR(sexdiff_full_lm_combat)
FDR_sexdiff_full_lm_combat
```
## Calculate t and p values
```{r}
dof <- 424
tvals_sex <- abs(sexdiff_full_lm_combat[, 'tvalue-Mouse_SexM'])
pvalsex <- 2*pt(q = tvals_sex, df = dof, lower.tail = FALSE) %>% as_tibble()

tvals_ttv <- abs(sexdiff_full_lm_combat[, 'tvalue-total_tissue_volume'])
pvalttv <- 2*pt(q = tvals_ttv, df = dof, lower.tail = FALSE) %>% as_tibble()

tvals_age <- abs(sexdiff_full_lm_combat[, 'tvalue-AGE_cent'])
pvalage <- 2*pt(q = tvals_age, df = dof, lower.tail = FALSE) %>% as_tibble()

model_outputs_sex_wTTV_cobmat <- sexdiff_full_lm_combat %>% as.data.frame()
model_outputs_sex_wTTV_cobmat$label <- rownames(sexdiff_full_lm_combat)
model_outputs_sex_wTTV1_cobmat <- cbind(model_outputs_sex_wTTV_cobmat[c("label","F-statistic","R-squared","beta-(Intercept)","beta-Mouse_SexM","beta-AGE_cent",
                                                          "beta-total_tissue_volume","tvalue-(Intercept)","tvalue-Mouse_SexM","tvalue-AGE_cent" , 
                                                          "tvalue-total_tissue_volume")], 
                                 pvalsex,pvalttv,pvalage,FDR_sexdiff_full_lm_combat[,c("qvalue-tvalue-Mouse_SexM","qvalue-tvalue-AGE_cent",
                                 "qvalue-tvalue-total_tissue_volume")])

colnames(model_outputs_sex_wTTV1_cobmat) <- c("label","F-statistic", "R-squared", "beta-(Intercept)", "beta-Mouse_SexM","beta-AGE_cent", "beta-total_tissue_volume", 
                                       "tvalue-(Intercept)", "tvalue-Mouse_SexM", "tvalue-AGE_cent","tvalue-total_tissue_volume", 
                                       "pvalsex","pvalttv","pvalage",
                                       "qvalue-tvalue-Mouse_SexM","qvalue-tvalue-AGE_cent","qvalue-tvalue-total_tissue_volume")
```

### Plot results
Threshold regions that are significant at 5% FDR (t=2.29) and 20% FDR (t=1.79) 
```{r}
## Threshold at p=0.05
df_beta_sex_wTTV_combat <- model_outputs_sex_wTTV1_cobmat[c("label", "beta-Mouse_SexM")]
df_tvalue_sex_wTTV_combat <- model_outputs_sex_wTTV1_cobmat[c("label", "tvalue-Mouse_SexM")]

## 5% FDR
sig_tvalue_sex_pos_wTBV_combat <- df_tvalue_sex_wTTV_combat$`tvalue-Mouse_SexM` > 2.19
sig_tvalue_sex_neg_wTBV_combat <- df_tvalue_sex_wTTV_combat$`tvalue-Mouse_SexM` < -2.19
sig_beta_sex_wTBV_combat <- df_beta_sex_wTTV_combat[sig_tvalue_sex_pos_wTBV_combat | sig_tvalue_sex_neg_wTBV_combat, ]

sig_beta_sex_tree_combat<-left_join(df_beta_sex_wTTV_combat, sig_beta_sex_wTBV_combat, by = "label")
filtered_sig_beta_sex_tree_combat <- subset(sig_beta_sex_tree_combat, select = c("label","beta-Mouse_SexM.y"))

hdefsCopy$Do(function(node){
  ind_5_combat <- which(filtered_sig_beta_sex_tree_combat$label == node$name)
  node$Sex_beta_5FDR_combat <- filtered_sig_beta_sex_tree_combat$`beta-Mouse_SexM.y`[ind_5_combat]
})

## 20% FDR
sig_tvalue_sex_pos20FDR_wTBV_combat <- df_tvalue_sex_wTTV_combat$`tvalue-Mouse_SexM` > 1.48
sig_tvalue_sex_neg20FDR_wTBV_combat <- df_tvalue_sex_wTTV_combat$`tvalue-Mouse_SexM` < -1.48
sig_beta_sex_20FDRwTBV_combat <- df_beta_sex_wTTV_combat[sig_tvalue_sex_pos20FDR_wTBV_combat | sig_tvalue_sex_neg20FDR_wTBV_combat, ]

sig_beta_sex_20FDRwTBV_tree_combat<-left_join(df_beta_sex_wTTV_combat, sig_beta_sex_20FDRwTBV_combat, by = "label")
filtered_sig_beta_sex_20FDRwTBV_tree_combat <- subset(sig_beta_sex_20FDRwTBV_tree_combat, select = c("label","beta-Mouse_SexM.y"))

hdefsCopy$Do(function(node){
  ind_20_combat <- which(filtered_sig_beta_sex_20FDRwTBV_tree_combat$label == node$name)
  node$Sex_beta_20FDR_combat <- filtered_sig_beta_sex_20FDRwTBV_tree_combat$`beta-Mouse_SexM.y`[ind_20_combat]
})

hdefsCopy$Do(function(node){
  ind_nothr_combat <- which(df_beta_sex_wTTV_combat$label == node$name)
  node$beta_sex_wTTV_combat <- df_beta_sex_wTTV_combat$`beta-Mouse_SexM`[ind_nothr_combat]
})
```


### Thresholded maps at 5%, 20%, and no threshold
```{r}
## load in beta values from the tree
sex_beta_wTBV_unthresholded_combat <- hanatToVolume(hdefsCopy, labelVol, "beta_sex_wTTV_combat")
sex_beta_wTBV_FDR05_combat <- hanatToVolume(hdefsCopy, labelVol, "Sex_beta_5FDR_combat") #turn the t-value back into a volume file
sex_beta_wTBV_FDR20_combat <- hanatToVolume(hdefsCopy, labelVol, "Sex_beta_20FDR_combat") #turn the t-value back into a volume file

#plot with no, 20% FDR, and 5%FDR thresholds
sliceSeries(nrow=6, ncol=2, begin=70, end=350) %>%
  anatomy(anatVol, low=700, high=1400) %>%
  overlay(sex_beta_wTBV_unthresholded_combat, low=0, high=1.5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("Unthresholded") %>% 
  sliceSeries() %>% anatomy() %>%
  overlay(sex_beta_wTBV_FDR20_combat, low=0, high=1.5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("20% FDR") %>% 
  sliceSeries() %>% anatomy() %>%
  overlay(sex_beta_wTBV_FDR05_combat, low=0, high=1.5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("5% FDR") %>% 
  legend("Standardized Effect Size (TTV corrected)") %>% 
  draw()

#plot with 5%FDR thresholds
sliceSeries(nrow=8, ncol=1, begin=70, end=350) %>%
  anatomy(anatVol, low=700, high=1400) %>%
  overlay(sex_beta_wTBV_unthresholded_combat, low=0, high=1.5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("Unthresholded") %>% 
  sliceSeries() %>% anatomy() %>%
  overlay(sex_beta_wTBV_FDR05_combat, low=0, high=1.5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("5% FDR") %>% 
  legend("Standardized Effect Size (TTV corrected)") %>% 
  draw()
```
 
## Density plot of beta values with TBV
```{r}
##With TBV
## select the columns of interest
df_mouse_results_wTBV <- model_outputs_sex_wTTV1_cobmat %>% dplyr::select(c(`label`, `beta-Mouse_SexM`))
## stack the columns containing the beta values for plotting
df_mouse_results_wTBV$Sexbias <- df_mouse_results_wTBV$`beta-Mouse_SexM` 
df_mouse_results_wTBV_1 <- df_mouse_results_wTBV %>% mutate(Sexbias = if_else(`Sexbias` > 0, "Male-bias", "Female-bias"))

## plot
ggplot(df_mouse_results_wTBV_1, aes(x=`beta-Mouse_SexM`, fill=`Sexbias`, color=`Sexbias`)) +
    geom_histogram(alpha=.4, binwidth = 0.05) +
    scale_fill_manual(values=c("Female-bias"="deepskyblue","Male-bias"="goldenrod1"))+
    scale_colour_manual(values=c("Female-bias"="deepskyblue","Male-bias"="goldenrod1"))+
    theme_classic()+ xlab("Standardized Effect Size (TTV corrected)")+ylab("Density")+
    scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) +
   theme(text = element_text(size = 14))+xlim(-1.5,1.5)+
    geom_vline(xintercept=c(0)) 

summary_table <- df_mouse_results_wTBV_1 %>% 
  dplyr::group_by(Sexbias) %>% 
  dplyr::summarize(
    mean_beta = mean(`beta-Mouse_SexM`),
    median_beta=median(`beta-Mouse_SexM`),
    sd_beta=sd(`beta-Mouse_SexM`),
    range_beta=range(`beta-Mouse_SexM`),
    n = n()
  )
summary_table

result = leveneTest(`beta-Mouse_SexM` ~ Sexbias, df_mouse_results_wTBV_1)
print(result)
```

#No TTV
## Lm for ComBat corrected vols
```{r}
sexdiff_full_lm_combat_noTTV <- anatLm(~ Mouse_Sex+Background+ AGE_cent, df_zscored_data_combat, df_zscored_data_combat[,13:ncol(df_zscored_data_combat)])
FDR_sexdiff_full_lm_combat_noTTV <- anatFDR(sexdiff_full_lm_combat_noTTV)
FDR_sexdiff_full_lm_combat_noTTV
```

### Plot results t-values for Background
Threshold regions that are significant at 5% FDR (t=2.49) and 20% FDR
```{r}
model_outputs_sexdiff_full_lm_combat_noTTV <- sexdiff_full_lm_combat_noTTV %>% as.data.frame()
model_outputs_sexdiff_full_lm_combat_noTTV$label <- rownames(sexdiff_full_lm_combat_noTTV)

hdefsCopy$Do(function(node){
  ind_background_noTTV <- which(model_outputs_sexdiff_full_lm_combat_noTTV$label == node$name)
  node$background_noTTV <- model_outputs_sexdiff_full_lm_combat_noTTV$`tvalue-BackgroundC57BL-6N`[ind_background_noTTV]
})
t_value_background_noTTV <- hanatToVolume(hdefsCopy, labelVol, "background_noTTV")

sliceSeries(nrow=6, ncol=2, begin=70, end=350) %>%
  anatomy(anatVol, low=700, high=1400) %>%
  overlay(t_value_background_noTTV, low=2.10, high=5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("Background Strain 5%FDR") %>% 
  legend("T-value (not TTV corrected)") %>% 
  draw()
```

### Plot results t-values
Threshold regions that are significant at 5% FDR (t=2.49) and 20% FDR
```{r}
hdefsCopy$Do(function(node){
  ind_sex_full_model_combat_noTTV <- which(model_outputs_sexdiff_full_lm_combat_noTTV$label == node$name)
  node$sex_full_model_combat_noTTV <- model_outputs_sexdiff_full_lm_combat_noTTV$`tvalue-Mouse_SexM`[ind_sex_full_model_combat_noTTV]
})
t_value_sex_full_model_combat_noTTV <- hanatToVolume(hdefsCopy, labelVol, "sex_full_model_combat_noTTV")

sliceSeries(nrow=6, ncol=2, begin=70, end=350) %>%
  anatomy(anatVol, low=700, high=1400) %>%
  overlay(t_value_sex_full_model_combat_noTTV, low=2.34, high=5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("Sex 5%FDR") %>% 
  legend("T-value (not TTV corrected)") %>% 
  draw()
```



```{r}
dof <- 424
tvals_sex_noTTV <- abs(sexdiff_full_lm_combat_noTTV[, 'tvalue-Mouse_SexM'])
pvalsex_noTTV <- 2*pt(q = tvals_sex, df = dof, lower.tail = FALSE) %>% as_tibble()

tvals_age_noTTV <- abs(sexdiff_full_lm_combat_noTTV[, 'tvalue-AGE_cent'])
pvalage_noTTV <- 2*pt(q = tvals_age, df = dof, lower.tail = FALSE) %>% as_tibble()

model_outputs_sex_wTTV_cobmat_noTTV <- sexdiff_full_lm_combat_noTTV %>% as.data.frame()
model_outputs_sex_wTTV_cobmat_noTTV$label <- rownames(sexdiff_full_lm_combat_noTTV)
model_outputs_sex_wTTV1_cobmat_noTTV <- cbind(model_outputs_sex_wTTV_cobmat_noTTV[c("label","F-statistic","R-squared","beta-(Intercept)","beta-Mouse_SexM","beta-AGE_cent",
                                                          "tvalue-(Intercept)","tvalue-Mouse_SexM","tvalue-AGE_cent")], 
                                 pvalsex_noTTV,pvalage_noTTV,FDR_sexdiff_full_lm_combat_noTTV[,c("qvalue-tvalue-Mouse_SexM","qvalue-tvalue-AGE_cent")])

colnames(model_outputs_sex_wTTV1_cobmat_noTTV) <- c("label","F-statistic", "R-squared", "beta-(Intercept)", "beta-Mouse_SexM","beta-AGE_cent",  
                                       "tvalue-(Intercept)", "tvalue-Mouse_SexM", "tvalue-AGE_cent",
                                       "pvalsex","pvalage","qvalue-tvalue-Mouse_SexM","qvalue-tvalue-AGE_cent")
```

### Plot results
Threshold regions that are significant at 5% FDR (t=2.29) and 20% FDR (t=1.79) 
```{r}
## Threshold at p=0.05
df_beta_sex_wTTV_combat_noTTV <- model_outputs_sex_wTTV1_cobmat_noTTV[c("label", "beta-Mouse_SexM")]
df_tvalue_sex_wTTV_combat_noTTV <- model_outputs_sex_wTTV1_cobmat_noTTV[c("label", "tvalue-Mouse_SexM")]

## 5% FDR
sig_tvalue_sex_pos_wTBV_combat_noTTV <- df_tvalue_sex_wTTV_combat_noTTV$`tvalue-Mouse_SexM` > 2.32
sig_tvalue_sex_neg_wTBV_combat_noTTV <- df_tvalue_sex_wTTV_combat_noTTV$`tvalue-Mouse_SexM` < -2.32
sig_beta_sex_wTBV_combat_noTTV <- df_beta_sex_wTTV_combat_noTTV[sig_tvalue_sex_pos_wTBV_combat_noTTV | sig_tvalue_sex_neg_wTBV_combat_noTTV, ]

sig_beta_sex_tree_combat_noTTV<-left_join(df_beta_sex_wTTV_combat_noTTV, sig_beta_sex_wTBV_combat_noTTV, by = "label")
filtered_sig_beta_sex_tree_combat_noTTV <- subset(sig_beta_sex_tree_combat_noTTV, select = c("label","beta-Mouse_SexM.y"))

hdefsCopy$Do(function(node){
  ind_5_combat_noTTV <- which(filtered_sig_beta_sex_tree_combat_noTTV$label == node$name)
  node$Sex_beta_5FDR_combat_noTTV <- filtered_sig_beta_sex_tree_combat_noTTV$`beta-Mouse_SexM.y`[ind_5_combat_noTTV]
})

## 20% FDR
sig_tvalue_sex_pos20FDR_wTBV_combat_noTTV <- df_tvalue_sex_wTTV_combat_noTTV$`tvalue-Mouse_SexM` > 1.55
sig_tvalue_sex_neg20FDR_wTBV_combat_noTTV <- df_tvalue_sex_wTTV_combat_noTTV$`tvalue-Mouse_SexM` < -1.55
sig_beta_sex_20FDRwTBV_combat_noTTV <- df_beta_sex_wTTV_combat_noTTV[sig_tvalue_sex_pos20FDR_wTBV_combat_noTTV | sig_tvalue_sex_neg20FDR_wTBV_combat_noTTV, ]

sig_beta_sex_20FDRwTBV_tree_combat_noTTV<-left_join(df_beta_sex_wTTV_combat_noTTV, sig_beta_sex_20FDRwTBV_combat_noTTV, by = "label")
filtered_sig_beta_sex_20FDRwTBV_tree_combat_noTTV <- subset(sig_beta_sex_20FDRwTBV_tree_combat_noTTV, select = c("label","beta-Mouse_SexM.y"))

hdefsCopy$Do(function(node){
  ind_20_combat_noTTV <- which(filtered_sig_beta_sex_20FDRwTBV_tree_combat_noTTV$label == node$name)
  node$Sex_beta_20FDR_combat_noTTV <- filtered_sig_beta_sex_20FDRwTBV_tree_combat_noTTV$`beta-Mouse_SexM.y`[ind_20_combat_noTTV]
})

hdefsCopy$Do(function(node){
  ind_nothr_combat_noTTV <- which(df_beta_sex_wTTV_combat_noTTV$label == node$name)
  node$beta_sex_wTTV_combat_noTTV <- df_beta_sex_wTTV_combat_noTTV$`beta-Mouse_SexM`[ind_nothr_combat_noTTV]
})
```


### Thresholded maps at 5%, 20%, and no threshold
```{r}
## load in beta values from the tree
sex_beta_wTBV_unthresholded_combat_noTTV <- hanatToVolume(hdefsCopy, labelVol, "beta_sex_wTTV_combat_noTTV")
sex_beta_wTBV_FDR05_combat_noTTV <- hanatToVolume(hdefsCopy, labelVol, "Sex_beta_5FDR_combat_noTTV") #turn the t-value back into a volume file
sex_beta_wTBV_FDR20_combat_noTTV <- hanatToVolume(hdefsCopy, labelVol, "Sex_beta_20FDR_combat_noTTV") #turn the t-value back into a volume file

sliceSeries(nrow=6, ncol=2, begin=70, end=350) %>%
  anatomy(anatVol, low=700, high=1400) %>%
  overlay(sex_beta_wTBV_unthresholded_combat_noTTV, low=0, high=1.5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("Unthresholded") %>% 
  sliceSeries() %>% anatomy() %>%
  overlay(sex_beta_wTBV_FDR20_combat_noTTV, low=0, high=1.5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("20% FDR") %>% 
  sliceSeries() %>% anatomy() %>%
  overlay(sex_beta_wTBV_FDR05_combat_noTTV, low=0, high=1.5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("5% FDR") %>% 
  legend("Standardized Effect Size (not TTV corrected)") %>% 
  draw()

sliceSeries(nrow=8, ncol=1, begin=70, end=350) %>%
  anatomy(anatVol, low=700, high=1400) %>%
  overlay(sex_beta_wTBV_unthresholded_combat_noTTV, low=0, high=1.5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("Unthresholded") %>% 
  sliceSeries() %>% anatomy() %>%
  overlay(sex_beta_wTBV_FDR05_combat_noTTV, low=0, high=1.5, col = positive_cols(), rCol=negative_cols(), symmetric = T) %>% addtitle("5% FDR") %>% 
  legend("Standardized Effect Size (not TTV corrected)") %>% 
  draw()
```


## Density plot of beta values without TBV
```{r}
## select the columns of interest
df_mouse_results_wTBV_noTTV <- model_outputs_sexdiff_full_lm_combat_noTTV %>% dplyr::select(c(`label`, `beta-Mouse_SexM`))
## stack the columns containing the beta values for plotting
df_mouse_results_wTBV_noTTV$Sexbias <- df_mouse_results_wTBV_noTTV$`beta-Mouse_SexM` 
df_mouse_results_wTBV_1_noTTV <- df_mouse_results_wTBV_noTTV %>% mutate(Sexbias = if_else(`Sexbias` > 0, "Male-bias", "Female-bias"))
## plot
ggplot(df_mouse_results_wTBV_1_noTTV, aes(x=`beta-Mouse_SexM`, fill=`Sexbias`, color=`Sexbias`)) +
    geom_histogram(alpha=.4, binwidth = 0.05) +
    scale_fill_manual(values=c("Female-bias"="deepskyblue","Male-bias"="goldenrod1"))+
    scale_colour_manual(values=c("Female-bias"="deepskyblue","Male-bias"="goldenrod1"))+
    theme_classic()+ xlab("Standardized Effect Size (TTV corrected)")+ylab("Density")+
    scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) +
   theme(text = element_text(size = 14))+xlim(-1.5,1.5)+
    geom_vline(xintercept=c(0)) 

summary_table_noTTV <- df_mouse_results_wTBV_1_noTTV %>% 
  dplyr::group_by(Sexbias) %>% 
  dplyr::summarize(
    median_beta=median(`beta-Mouse_SexM`),
    mean_beta = mean(`beta-Mouse_SexM`),
    sd_beta=sd(`beta-Mouse_SexM`),
    range_beta=range(`beta-Mouse_SexM`),
    n = n()
  )
summary_table_noTTV
```


# Variance analyses
Residulaize volumes & look at variance with TTV
```{r}
## Calculate residualized volumes for males
df_just_males <-df_zscored_data_combat %>% filter(Mouse_Sex == "M")
mouse_vols_males <- df_just_males[,13:ncol(df_just_males)] #Then subset the volumes for the control observations #vols from 54:461
mouse_demo_males <- df_just_males[,1:12]

residuals_mouse_males <- matrix(0, nrow = nrow(mouse_vols_males), ncol = ncol(mouse_vols_males))
for(j in 1:ncol(mouse_vols_males)){
  residuals_mouse_males[,j] <- residuals(lm(mouse_vols_males[,j] ~ Background + AGE_cent + total_tissue_volume, data=df_just_males))
}
### Assign column names and "rebind" demographics data
colnames(residuals_mouse_males)<-colnames(mouse_vols_males)
df_residuals_mouse_males <- mouse_demo_males %>% cbind(residuals_mouse_males)

## Calculate residualized volumes for females
df_just_females <-df_zscored_data_combat %>% filter(Mouse_Sex == "F")
mouse_vols_females <- df_just_females[,13:ncol(df_just_females)] #Then subset the volumes for the control observations #vols from 54:461
mouse_demo_females <- df_just_females[,1:12]

residuals_mouse_females <- matrix(0, nrow = nrow(mouse_vols_females), ncol = ncol(mouse_vols_females))
for(j in 1:ncol(mouse_vols_females)){
  residuals_mouse_females[,j] <- residuals(lm(mouse_vols_females[,j] ~ Background + AGE_cent + total_tissue_volume, data=df_just_females))
}
### Assign column names and "rebind" demographics data
colnames(residuals_mouse_females)<-colnames(mouse_vols_females)
df_residuals_mouse_females <- mouse_demo_females %>% cbind(residuals_mouse_females)

df_residuals_mouse <- rbind(df_residuals_mouse_males, df_residuals_mouse_females)
residuals_mouse <- df_residuals_mouse[,13:ncol(df_residuals_mouse)]

## Calculate variance
variance_males_mouse <- matrix(0, nrow = 1, ncol = ncol(residuals_mouse_males))
for(j in 1:ncol(residuals_mouse_males)){
  variance_males_mouse[,j] <- var(residuals_mouse_males[,j]) #Then you can compute the ROI-wise averages
}
colnames(variance_males_mouse)<-colnames(residuals_mouse_males)

variance_females_mouse <- matrix(0, nrow = 1, ncol = ncol(residuals_mouse_females))
for(j in 1:ncol(residuals_mouse_females)){
  variance_females_mouse[,j] <- var(residuals_mouse_females[,j]) #Then you can compute the ROI-wise averages
}
colnames(variance_females_mouse)<-colnames(residuals_mouse_females)

variance_mouse <- t(rbind(variance_males_mouse, variance_females_mouse)) %>% as.data.frame()
colnames(variance_mouse) <- c("male_variance", "female_variance")
variance_mouse_1 <- variance_mouse %>% mutate(Sexbias = if_else(male_variance > female_variance, "Male-bias", "Female-bias"))
variance_mouse_1$label <- rownames(variance_mouse_1)

## Perform Levene's test across all regions
Levene_test_pvalue <- matrix(NA, nrow = ncol(residuals_mouse), ncol= 1)
Levene_test_fvalue <- matrix(NA, nrow = ncol(residuals_mouse), ncol= 1)

for(i in 1:ncol(residuals_mouse)){
    result <- leveneTest(residuals_mouse[,i] ~ Mouse_Sex, df_residuals_mouse)
#store outputs of interest
Levene_test_pvalue[i,] <- result[1,3] # extract Pr(>F)
Levene_test_fvalue[i,] <- result[1,2] # extract F value
}

variance_analysis_mice<- cbind(Levene_test_fvalue, Levene_test_pvalue) %>% as.data.frame()
variance_analysis_mice$label <- colnames(residuals_mouse)
colnames(variance_analysis_mice) <- c("fvalue","pvalue","label")
variance_analysis_mice$Hochberg <- p.adjust(variance_analysis_mice$pvalue,method = "hochberg")
variance_analysis_mouse_1 <- left_join(variance_analysis_mice, variance_mouse_1, by="label")

variance_analysis_mouse_2 <- read.csv("/Users/gumae2/Documents/Cross_species_sex_differences/Tables/Variance_analysis_mouse_2.csv")
## p<0.05
sig_positive_pvalues <- variance_analysis_mouse_2$pvalue < 0.049
sig_positive_pvalues_levenes <- variance_analysis_mouse_2[sig_positive_pvalues, ]

df_sig<-left_join(variance_analysis_mouse_2, sig_positive_pvalues_levenes, by = "label")
filtered_df_sig <- subset(df_sig, select = c("label","sex.biased.pvalue.y"))

hdefsCopy$Do(function(node){
  ind_pvalue_levene <- which(filtered_df_sig$label == node$name)
  node$Pvalue_levene <- filtered_df_sig$`sex.biased.pvalue.y`[ind_pvalue_levene]
})

pvalue_levenes <- hanatToVolume(hdefsCopy, labelVol, "Pvalue_levene")

### Assign colour palette for positive and negative beta values 

positive_cols_variance <- function() {
  getOption("MRIcrotomeCol", colorRampPalette(c("purple","violet"))(255)) }
negative_cols_variance <- function() {
  getOption("MRIcrotomeCol", colorRampPalette(c("green","light yellow"))(255))
}

sliceSeries(nrow=6, ncol=2, begin=70, end=350) %>%
  anatomy(anatVol, low=700, high=1400) %>%
  overlay(pvalue_levenes, low=0, high=0.05, col = positive_cols_variance(), rCol= negative_cols_variance(), symmetric = T) %>% addtitle("p<0.05") %>% 
  legend("p-value (Levene's)") %>% 
  draw()
```

# Variance analyses (no TTV correction)
Residulaize volumes & look at variance without TTV
```{r}
## Calculate residualized volumes for males
df_just_males_noTTV <-df_zscored_data_combat %>% filter(Mouse_Sex == "M")
mouse_vols_males_noTTV <- df_just_males_noTTV[,13:ncol(df_just_males_noTTV)] #Then subset the volumes for the control observations #vols from 54:461
mouse_demo_males_noTTV <- df_just_males_noTTV[,1:12]

residuals_mouse_males_noTTV <- matrix(0, nrow = nrow(mouse_vols_males_noTTV), ncol = ncol(mouse_vols_males_noTTV))
for(j in 1:ncol(mouse_vols_males_noTTV)){
  residuals_mouse_males_noTTV[,j] <- residuals(lm(mouse_vols_males_noTTV[,j] ~ Background + AGE_cent, data=df_just_males_noTTV))
}
### Assign column names and "rebind" demographics data
colnames(residuals_mouse_males_noTTV)<-colnames(mouse_vols_males_noTTV)
df_residuals_mouse_males_noTTV <- mouse_demo_males_noTTV %>% cbind(residuals_mouse_males_noTTV)

## Calculate residualized volumes for females
df_just_females_noTTV <-df_zscored_data_combat %>% filter(Mouse_Sex == "F")
mouse_vols_females_noTTV <- df_just_females_noTTV[,13:ncol(df_just_females_noTTV)] #Then subset the volumes for the control observations #vols from 54:461
mouse_demo_females_noTTV <- df_just_females_noTTV[,1:12]

residuals_mouse_females_noTTV <- matrix(0, nrow = nrow(mouse_vols_females_noTTV), ncol = ncol(mouse_vols_females_noTTV))
for(j in 1:ncol(mouse_vols_females_noTTV)){
  residuals_mouse_females_noTTV[,j] <- residuals(lm(mouse_vols_females_noTTV[,j] ~ Background + AGE_cent, data=df_just_females_noTTV))
}
### Assign column names and "rebind" demographics data
colnames(residuals_mouse_females_noTTV)<-colnames(mouse_vols_females_noTTV)
df_residuals_mouse_females_noTTV <- mouse_demo_females_noTTV %>% cbind(residuals_mouse_females_noTTV)

df_residuals_mouse_noTTV <- rbind(df_residuals_mouse_males_noTTV, df_residuals_mouse_females_noTTV)
residuals_mouse_noTTV <- df_residuals_mouse_noTTV[,13:ncol(df_residuals_mouse_noTTV)]

## Calculate variance
variance_males_mouse_noTTV <- matrix(0, nrow = 1, ncol = ncol(residuals_mouse_males_noTTV))
for(j in 1:ncol(residuals_mouse_males_noTTV)){
  variance_males_mouse_noTTV[,j] <- var(residuals_mouse_males_noTTV[,j]) #Then you can compute the ROI-wise averages
}
colnames(variance_males_mouse_noTTV)<-colnames(residuals_mouse_males_noTTV)

variance_females_mouse_noTTV <- matrix(0, nrow = 1, ncol = ncol(residuals_mouse_females_noTTV))
for(j in 1:ncol(residuals_mouse_females_noTTV)){
  variance_females_mouse_noTTV[,j] <- var(residuals_mouse_females_noTTV[,j]) #Then you can compute the ROI-wise averages
}
colnames(variance_females_mouse_noTTV)<-colnames(residuals_mouse_females_noTTV)

variance_mouse_noTTV <- t(rbind(variance_males_mouse_noTTV, variance_females_mouse_noTTV)) %>% as.data.frame()
colnames(variance_mouse_noTTV) <- c("male_variance", "female_variance")
variance_mouse_1_noTTV <- variance_mouse_noTTV %>% mutate(Sexbias = if_else(male_variance > female_variance, "Male-bias", "Female-bias"))
variance_mouse_1_noTTV$label <- rownames(variance_mouse_1_noTTV)

## Perform Levene's test across all regions
Levene_test_pvalue_noTTV <- matrix(NA, nrow = ncol(residuals_mouse_noTTV), ncol= 1)
Levene_test_fvalue_noTTV <- matrix(NA, nrow = ncol(residuals_mouse_noTTV), ncol= 1)

for(i in 1:ncol(residuals_mouse_noTTV)){
    result_noTTV <- leveneTest(residuals_mouse_noTTV[,i] ~ Mouse_Sex, df_residuals_mouse_noTTV)
#store outputs of interest
Levene_test_pvalue_noTTV[i,] <- result_noTTV[1,3] # extract Pr(>F)
Levene_test_fvalue_noTTV[i,] <- result_noTTV[1,2] # extract F value
}

variance_analysis_mice_noTTV<- cbind(Levene_test_fvalue_noTTV, Levene_test_pvalue_noTTV) %>% as.data.frame()
variance_analysis_mice_noTTV$label <- colnames(residuals_mouse_noTTV)
colnames(variance_analysis_mice_noTTV) <- c("fvalue","pvalue_noTTV","label")
variance_analysis_mice_noTTV$Hochberg <- p.adjust(variance_analysis_mice_noTTV$pvalue_noTTV,method = "hochberg")
variance_analysis_mouse_1_noTTV <- left_join(variance_analysis_mice_noTTV, variance_mouse_1_noTTV, by="label")

variance_analysis_mouse_noTTV_2 <- read.csv("/Users/gumae2/Documents/Cross_species_sex_differences/Tables/Variance_analysis_mouse_noTTV_2.csv")

## p<0.05
sig_pvalues_noTTV <- variance_analysis_mouse_noTTV_2$pvalue_noTTV < 0.049
sig_pvalues_levenes_noTTV <- variance_analysis_mouse_noTTV_2[sig_pvalues_noTTV, ] %>% as.data.frame()

df_sig_noTTV <- left_join(variance_analysis_mice_noTTV, sig_pvalues_levenes_noTTV, by = "label")
filtered_df_sig_noTTV <- subset(df_sig_noTTV, select = c("label","sex.biased.pvalue"))

hdefsCopy$Do(function(node){
  ind_pvalue_levene_noTTV <- which(filtered_df_sig_noTTV$label == node$name)
  node$Pvalue_levene_noTTV <- filtered_df_sig_noTTV$`sex.biased.pvalue`[ind_pvalue_levene_noTTV]
})

pvalue_levenes_noTTV <- hanatToVolume(hdefsCopy, labelVol, "Pvalue_levene_noTTV")
### Assign colour palette for positive and negative beta values 

positive_cols_variance <- function() {
  getOption("MRIcrotomeCol", colorRampPalette(c("purple","violet"))(255)) }
negative_cols_variance <- function() {
  getOption("MRIcrotomeCol", colorRampPalette(c("green","light yellow"))(255))
}

sliceSeries(nrow=6, ncol=2, begin=70, end=350) %>%
  anatomy(anatVol, low=700, high=1400) %>%
  overlay(pvalue_levenes_noTTV, low=0, high=0.05, col = positive_cols_variance(), rCol= negative_cols_variance(), symmetric = T) %>% addtitle("p<0.05") %>% 
  legend("p-value (Levene's)") %>% 
  draw()
```









