---
title: "Sex-diffs-homologous-ROIs-fullmouse-humans"
author: "Elisa Guma"
date: "2/27/2023"
output: html_document
---
# Description

In this script, we identify a subset of brain regions deemed to be "homologous" across humans and mice. 
We pull these regions (and aggregate across some smaller ones) to obtain a final list of volumes. 
We use these to correlate the effects of sex on brain volume across species. 

# Load in necessary libraries
```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RMINC)
library(broom)
library(data.tree)
library(ggrepel)
library(ggplot2)
library(readr)
library(sva)
library(WRS2)
source('../input_data/tree_tools.R')
```

# Import Human Data
```{r}
df_HCP_volumes <- readRDS("../input_data/human_anatomy/df_HCP_volumes_clean.RDS")
```

### Remove the subject with low euler number
```{r}
df_HCP_volumes_qc <- subset(df_HCP_volumes, df_HCP_volumes$euler > -200 & df_HCP_volumes$QC_include == 1)
```


## Select volumes for ROIs that will be used for homology analysis
Here we search for regions from our original data matrix that fall within our homologous region definition. ROI names come from the Glasser parcellation.
```{r}
HCP_data_homologs <- dplyr::select(df_HCP_volumes_qc, 
                           "SUBJECT_ID","Subject","Freesurfer_QC_07mm","AGE_cent","Age_in_Yrs","Sex","euler","BrainSegVolNotVent.y",
                           ends_with("Cerebellum.Cortex"),ends_with("Thalamus"),ends_with("Caudate"),
                           ends_with("Putamen"),ends_with("Pallidum"),ends_with("Whole_hippocampus"),
                           ends_with("CA1.body"), ends_with("CA1.head"), ends_with("CA3.body"), ends_with("CA3.head"), ends_with("GC.ML.DG.body"), ends_with("GC.ML.DG.head"),
                           ends_with("Whole_amygdala"),ends_with("Accumbens.area"), ends_with("Medial.nucleus"),
                           "Whole_brainstem","Midbrain","Pons","Medulla",
                           ends_with("BNST"),
                           ends_with("MPO"),
                           ends_with("LH"),
                           ends_with("VM"),
                           ends_with("DM"),
                           ends_with("AHA"),
                           ends_with("V1_ROI_volume"), 
                           ends_with("4_ROI_volume"), 
                           ends_with("3a_ROI_volume"), ends_with("3b_ROI_volume"), ends_with("1_ROI_volume"), ends_with("2_ROI_volume"), 
                           ends_with("A1_ROI_volume"),
                           ends_with("RSC_ROI_volume"),
                           ends_with("a24pr_ROI_volume"),ends_with("a24_ROI_volume"),ends_with("p24pr_ROI_volume"),
                           ends_with("p24_ROI_volume"), ends_with("24dd_ROI_volume"), ends_with("24dv_ROI_volume"),
                           ends_with("p32pr_ROI_volume"),ends_with("d32_ROI_volume"),ends_with("a32pr_ROI_volume"),
                           ends_with("p32_ROI_volume"),ends_with("s32_ROI_volume"),
                           ends_with("AVI_ROI_volume"), 
                           ends_with("AAIC_ROI_volume"), ends_with("MI_ROI_volume"),                           
                           ends_with("Pir_ROI_volume"),
                           ends_with("EC_ROI_volume"),
                           ends_with("PreS_ROI_volume"),
                           ends_with("PeEc_ROI_volume"),ends_with("TF_ROI_volume"),ends_with("PHA2_ROI_volume"), ends_with("PHA3_ROI_volume"),
                           ends_with("10r_ROI_volume"), ends_with("10v_ROI_volume"), 
                           ends_with("FFC_ROI_volume"), ends_with("PIT_ROI_volume"), ends_with("TE1a_ROI_volume"), ends_with("TE1p_ROI_volume"), 
                           ends_with("TE2a_ROI_volume"), ends_with("TF_ROI_volume"),
                           ends_with("STV_ROI_volume"), ends_with("STSvp_ROI_volume"), ends_with("STSva_ROI_volume"),
                           ends_with("5m_ROI_volume"), ends_with("5mv_ROI_volume"), ends_with("5L_ROI_volume")) 
```

#Aggregate data
```{r}
# Aggregate hypothalamus
HCP_data_homologs$`Left Hypothalamus` <- HCP_data_homologs$L_LH + HCP_data_homologs$L_VM + HCP_data_homologs$L_DM + HCP_data_homologs$L_AHA
HCP_data_homologs$`Right Hypothalamus` <- HCP_data_homologs$R_LH + HCP_data_homologs$R_VM + HCP_data_homologs$R_DM + HCP_data_homologs$R_AHA

# Aggregate primary somatosensory cortex
HCP_data_homologs$`Left Primary somatosensory area` <-HCP_data_homologs$`lh_L_3a_ROI_volume`+ HCP_data_homologs$`lh_L_3b_ROI_volume`+
  HCP_data_homologs$`lh_L_1_ROI_volume`+HCP_data_homologs$`lh_L_2_ROI_volume`
HCP_data_homologs$`Right Primary somatosensory area` <-HCP_data_homologs$`rh_R_3a_ROI_volume`+ HCP_data_homologs$`rh_R_3b_ROI_volume`+
  HCP_data_homologs$`rh_R_1_ROI_volume`+HCP_data_homologs$`rh_R_2_ROI_volume`

# Aggregate caudoputamen
HCP_data_homologs$`Left Caudoputamen` <-HCP_data_homologs$Left.Caudate+ HCP_data_homologs$Left.Putamen
HCP_data_homologs$`Right Caudoputamen` <-HCP_data_homologs$Right.Caudate+ HCP_data_homologs$Right.Putamen

# Aggregate CA1
HCP_data_homologs$`Left Field CA1` <-HCP_data_homologs$Left.CA1.head+ HCP_data_homologs$Left.CA1.head
HCP_data_homologs$`Right Field CA1` <-HCP_data_homologs$Right.CA1.head+ HCP_data_homologs$Right.CA1.head

# Aggregate CA3
HCP_data_homologs$`Left Field CA3` <-HCP_data_homologs$Left.CA3.body+ HCP_data_homologs$Left.CA3.head
HCP_data_homologs$`Right Field CA3` <-HCP_data_homologs$Right.CA3.body+ HCP_data_homologs$Right.CA3.head

# Aggregate DG
HCP_data_homologs$`Left Dentate gyrus, molecular layer` <-HCP_data_homologs$Left.GC.ML.DG.body+ HCP_data_homologs$Left.GC.ML.DG.head
HCP_data_homologs$`Right Dentate gyrus, molecular layer` <-HCP_data_homologs$Right.GC.ML.DG.body+ HCP_data_homologs$Right.GC.ML.DG.head

# Aggregate agranular insula cortex
HCP_data_homologs$`Left Agranular insula` <- HCP_data_homologs$lh_L_AVI_ROI_volume+
  HCP_data_homologs$lh_L_AAIC_ROI_volume+HCP_data_homologs$lh_L_MI_ROI_volume
HCP_data_homologs$`Right Agranular insula` <- HCP_data_homologs$rh_R_AVI_ROI_volume+
  HCP_data_homologs$rh_R_AAIC_ROI_volume+HCP_data_homologs$rh_R_MI_ROI_volume

# Aggregate anterior cingulate cortex
HCP_data_homologs$`Left Anterior cingulate area` <- HCP_data_homologs$lh_L_a24pr_ROI_volume+ HCP_data_homologs$lh_L_a24_ROI_volume+HCP_data_homologs$lh_L_p24pr_ROI_volume+
  HCP_data_homologs$lh_L_p24_ROI_volume+HCP_data_homologs$`lh_L_24dd_ROI_volume` +HCP_data_homologs$`lh_L_24dv_ROI_volume` +
  HCP_data_homologs$lh_L_p32pr_ROI_volume+ HCP_data_homologs$lh_L_d32_ROI_volume + HCP_data_homologs$lh_L_a32pr_ROI_volume + 
  HCP_data_homologs$lh_L_p32_ROI_volume + HCP_data_homologs$lh_L_s32_ROI_volume
HCP_data_homologs$`Right Anterior cingulate area` <- HCP_data_homologs$rh_R_a24pr_ROI_volume+ HCP_data_homologs$rh_R_a24_ROI_volume+HCP_data_homologs$rh_R_p24pr_ROI_volume+
  HCP_data_homologs$rh_R_p24_ROI_volume+HCP_data_homologs$`rh_R_24dd_ROI_volume` +HCP_data_homologs$`rh_R_24dv_ROI_volume` +
  HCP_data_homologs$rh_R_p32pr_ROI_volume+ HCP_data_homologs$rh_R_d32_ROI_volume + HCP_data_homologs$rh_R_a32pr_ROI_volume + 
  HCP_data_homologs$rh_R_p32_ROI_volume + HCP_data_homologs$rh_R_s32_ROI_volume

# Aggregate perirhinal cortex                            
HCP_data_homologs$`Left Perirhinal area` <- HCP_data_homologs$lh_L_PeEc_ROI_volume +HCP_data_homologs$lh_L_TF_ROI_volume+ HCP_data_homologs$lh_L_PHA2_ROI_volume+
  HCP_data_homologs$lh_L_PHA3_ROI_volume
HCP_data_homologs$`Right Perirhinal area` <- HCP_data_homologs$rh_R_PeEc_ROI_volume +HCP_data_homologs$rh_R_TF_ROI_volume+ HCP_data_homologs$rh_R_PHA2_ROI_volume+
  HCP_data_homologs$rh_R_PHA3_ROI_volume

# Aggregate ventral orbital area
HCP_data_homologs$`Left Ventral orbital area` <- HCP_data_homologs$`lh_L_10r_ROI_volume` +HCP_data_homologs$`lh_L_10v_ROI_volume`
HCP_data_homologs$`Right Ventral orbital area` <- HCP_data_homologs$`rh_R_10r_ROI_volume` +HCP_data_homologs$`rh_R_10v_ROI_volume`

# Aggregate temporal association cortex
HCP_data_homologs$`Left Temporal association areas` <- HCP_data_homologs$lh_L_FFC_ROI_volume+ HCP_data_homologs$lh_L_PIT_ROI_volume+ HCP_data_homologs$lh_L_TE1a_ROI_volume+ 
HCP_data_homologs$lh_L_TE1p_ROI_volume+ HCP_data_homologs$lh_L_TE2a_ROI_volume+ HCP_data_homologs$lh_L_TF_ROI_volume+HCP_data_homologs$lh_L_STV_ROI_volume+ HCP_data_homologs$lh_L_STSvp_ROI_volume+ HCP_data_homologs$lh_L_STSva_ROI_volume
HCP_data_homologs$`Right Temporal association areas` <- HCP_data_homologs$rh_R_FFC_ROI_volume+ HCP_data_homologs$rh_R_PIT_ROI_volume+ HCP_data_homologs$rh_R_TE1a_ROI_volume+ 
HCP_data_homologs$rh_R_TE1p_ROI_volume+ HCP_data_homologs$rh_R_TE2a_ROI_volume+ HCP_data_homologs$rh_R_TF_ROI_volume+HCP_data_homologs$rh_R_STV_ROI_volume+ HCP_data_homologs$rh_R_STSvp_ROI_volume+ HCP_data_homologs$rh_R_STSva_ROI_volume

#Aggregate posterior parietal cortex
HCP_data_homologs$`Left Posterior parietal association areas` <- HCP_data_homologs$`lh_L_5m_ROI_volume`+ HCP_data_homologs$`lh_L_5mv_ROI_volume`+ HCP_data_homologs$`lh_L_5L_ROI_volume`
HCP_data_homologs$`Right Posterior parietal association areas` <- HCP_data_homologs$`rh_R_5m_ROI_volume`+ HCP_data_homologs$`rh_R_5mv_ROI_volume`+ HCP_data_homologs$`rh_R_5L_ROI_volume`

# Rename regions so that they match the mouse names later on
HCP_data_homologs <-dplyr::rename(HCP_data_homologs,
                          `Left Primary visual area`=lh_L_V1_ROI_volume,`Right Primary visual area`=rh_R_V1_ROI_volume,
                          `Left Primary motor area` = `lh_L_4_ROI_volume`,`Right Primary motor area` = `rh_R_4_ROI_volume`, 
                          `Left Primary auditory area` = lh_L_A1_ROI_volume, `Right Primary auditory area` = rh_R_A1_ROI_volume,
                          `Left Retrosplenial area`=lh_L_RSC_ROI_volume, `Right Retrosplenial area`=rh_R_RSC_ROI_volume,
                          `Left Piriform cortex` =lh_L_Pir_ROI_volume,`Right Piriform cortex` =rh_R_Pir_ROI_volume,
                          `Left Entorhinal area` =lh_L_EC_ROI_volume,`Right Entorhinal area` =rh_R_EC_ROI_volume,
                          `Left Subiculum`=lh_L_PreS_ROI_volume,`Right Subiculum`=rh_R_PreS_ROI_volume,
                          `Left Cerebellar cortex`=Left.Cerebellum.Cortex,`Right Cerebellar cortex`=Right.Cerebellum.Cortex,
                          `Left Nucleus accumbens`=Left.Accumbens.area,`Right Nucleus accumbens`=Right.Accumbens.area,
                          `Left Globus pallidus`=`Left.Pallidum`,`Right Globus pallidus`=`Right.Pallidum`,
                          `Left Amygdala` = `Left.Whole_amygdala`, `Right Amygdala` = `Right.Whole_amygdala`,
                          `Left Hippocampus` = `Left.Whole_hippocampus`, `Right Hippocampus` = `Right.Whole_hippocampus`,
                          `Left Medial amygdalar nucleus` = `Left.Medial.nucleus`, `Right Medial amygdalar nucleus` = `Right.Medial.nucleus`, 
                          `Left Thalamus` = `Left.Thalamus`, `Right Thalamus` = `Right.Thalamus`,
                          `Brain stem`=`Whole_brainstem`, 
                          `Left Bed nuclei of the stria terminalis` = L_BNST, `Right Bed nuclei of the stria terminalis` = R_BNST, 
                          `Left Medial preoptic nucleus` = L_MPO, `Right Medial preoptic nucleus` = R_MPO)


# Clean up the names
human_HCP_data_homologs <- dplyr::select(HCP_data_homologs, -ends_with("volume"), -"Left.Caudate",-"Left.Putamen", -"Right.Caudate",-"Right.Putamen",
                                         -ends_with("body"), -ends_with("head"), -ends_with("LH"), -ends_with("VM"), -ends_with("DM"),- ends_with("AHA"))
```

## Generate weights for the gene aggregation
```{r}
HCP_data_homologs_vols <- human_HCP_data_homologs[,8:ncol(human_HCP_data_homologs)]
mean_vols <- colMeans(HCP_data_homologs_vols) %>% as.data.frame()
colnames(mean_vols) <- "mean_vol_agg"
mean_vols$label <- rownames(mean_vols)
mean_vols$aggregate_label <- mean_vols$label
#write.csv(mean_vols, file="aggregated_vols_human.csv")
```

## Zscore Human Data
```{r}
human_vols <- human_HCP_data_homologs[,8:ncol(human_HCP_data_homologs)] 
HCP_demo <- human_HCP_data_homologs[,1:7]

zscores_HCP <- matrix(0, nrow = nrow(human_vols), ncol = ncol(human_vols))
for(j in 1:ncol(human_vols)){
  mu <- mean(human_vols[,j]) #Then you can compute the ROI-wise averages
  sigma <- sd(human_vols[,j]) #And standard deviations
  zscores_HCP[,j] <- (human_vols[,j] - mu)/sigma #Then you can use those to make the z-score volumes
}

### Assign column names and "rebind" demographics data
colnames(zscores_HCP)<-colnames(human_vols)
df_HCP_homologs <- HCP_demo %>% cbind(zscores_HCP) 
```


### With TBV covariate
Here we run a linear model testing for group differences (XXY vs XY) covarying for age and total tissue volume so that we can extract standardized beta values for each region of interest.
```{r}
Sex_diffs_homologs_lm_wTBV= anatLm(~ Sex + AGE_cent + BrainSegVolNotVent.y + euler, df_HCP_homologs,
                             df_HCP_homologs[,8:ncol(df_HCP_homologs)])
FDR_Sex_Lm_wTBV_zscore <- anatFDR(Sex_diffs_homologs_lm_wTBV)
FDR_Sex_Lm_wTBV_zscore

Human_sex_model_outputs_wTBV <- Sex_diffs_homologs_lm_wTBV %>% as.data.frame()
Human_sex_model_outputs_wTBV$label <- rownames(Human_sex_model_outputs_wTBV)
```

### Without TBV covariate
XXY (repeat the above analysis without total tissue volume correction)
```{r}
Sex_diffs_homologs_lm_noTBV= anatLm(~ Sex + AGE_cent + euler, df_HCP_homologs,
                             df_HCP_homologs[,8:ncol(df_HCP_homologs)])
FDR_Sex_Lm_noTBV_zscore <- anatFDR(Sex_diffs_homologs_lm_noTBV)
FDR_Sex_Lm_noTBV_zscore

Human_sex_model_outputs_noTBV <- Sex_diffs_homologs_lm_noTBV %>% as.data.frame()
Human_sex_model_outputs_noTBV$label <- rownames(Human_sex_model_outputs_noTBV)
```


# Import Mouse Data
```{r}
gf <- read_csv("../input_data/mouse_anatomy/mouse_demographics.csv")
treeVols <- readRDS("../input_data/mouse_anatomy/full_mouse_sampletree.RDS")
source('../input_data/tree_tools.R')
```

##First prune out the ventricles and white matter from the tree
```{r}
#Create a copy of the tree
tree_pruned <- Clone(treeVols)
nodes_to_cut <- c("ventricular systems", "fiber tracts", "medial forebrain bundle system",
                  "lateral forebrain bundle system","extrapyramidal fiber systems") #remove ventricles and cranial nerves
pruneAnatTree(tree_pruned, nodes = nodes_to_cut, method = "AtNode")
```

## Prune tree to have only left ROIs and re-aggregate up the tree
```{r}
left_tree <- Clone(tree_pruned)
Prune(left_tree, pruneFun = function(node){!str_detect(node$name, "right")})
left_tree_clean <- Clone(left_tree)
nodes_to_cut <- c("Vermal regions", "Midbrain", "Hindbrain") #remove midline regions
pruneAnatTree(left_tree_clean, nodes = nodes_to_cut, method = "AtNode")

#Re-aggregate the volumes from the left leaf nodes up the tree
left_tree_clean$Do(function(node){
  node$volumes <- Aggregate(node, 'volumes', rowSums)
}, traversal = 'post-order')
```

## Prune tree to have only right ROIs & reaggregate
```{r}
right_tree <- Clone(tree_pruned)
Prune(right_tree, pruneFun = function(node){!str_detect(node$name, "left")})
right_tree_clean <- Clone(right_tree)
nodes_to_cut <- c("Vermal regions", "Midbrain", "Hindbrain") #remove midline regions
pruneAnatTree(right_tree_clean, nodes = nodes_to_cut, method = "AtNode")

#Re-aggregate the volumes from the right leaf nodes up the tree
right_tree_clean$Do(function(node){
  node$volumes <- Aggregate(node, 'volumes', rowSums)
}, traversal = 'post-order')
```

## Prune midline tree to have only right ROIs & reaggregate
```{r}
midline_tree <- Clone(tree_pruned)
Prune(midline_tree, pruneFun = function(node){!str_detect(node$name, "left|right")})
midline_tree_clean <- Clone(midline_tree)
nodes_to_merge <- c("Vermal regions", "Midbrain", "Medulla", "Pons") #select midline regions
pruneAnatTree(midline_tree_clean, nodes = nodes_to_merge, method = "BelowNode")

#Re-aggregate the volumes from the right leaf nodes up the tree
midline_tree_clean$Do(function(node){
  node$volumes <- Aggregate(node, 'volumes', rowSums)
}, traversal = 'post-order')
```

## Aggregate for the left
```{r}
# prune nodes of interest:
left_tree_for_homology <- Clone(left_tree_clean)
nodes_to_merge<- c("Piriform cortex", 
                   "Subiculum",
                   "Entorhinal area",
                   "Field CA1",
                   "Field CA3",
                   "Dentate gyrus, molecular layer",
                   "Medial amygdalar nucleus",
                   "Cortical subplate",
                   "Anterior cingulate area",
                   "Retrosplenial area",
                   "Primary auditory area",
                   "Agranular insular area",
                   "Perirhinal area",
                   "Primary motor area",
                   "Orbital area, ventrolateral part",
                   "Primary somatosensory area",
                   "Primary visual area",
                   "Temporal association areas",
                   "Posterior parietal association areas",
                   "Striatum ventral region",
                   "Caudoputamen",
                   "Thalamus",
                   "Medial preoptic nucleus",
                   "Hemispheric regions")
pruneAnatTree(left_tree_for_homology, nodes = nodes_to_merge, method = "BelowNode")

left_tree_leaves <- left_tree_for_homology$Get("name", filterFun = isLeaf) %>% as.data.frame()
```
## Aggregate for the right
```{r}
# prune nodes of interest:
right_tree_for_homology <- Clone(right_tree_clean)
nodes_to_merge<- c("Piriform cortex", 
                   "Subiculum",
                   "Entorhinal area",
                   "Field CA1",
                   "Field CA3",
                   "Dentate gyrus, molecular layer",
                   "Medial amygdalar nucleus",
                   "Cortical subplate",
                   "Anterior cingulate area",
                   "Retrosplenial area",
                   "Primary auditory area",
                   "Agranular insular area",
                   "Perirhinal area",
                   "Primary motor area",
                   "Orbital area, ventrolateral part",
                   "Primary somatosensory area",
                   "Primary visual area",
                   "Temporal association areas",
                   "Posterior parietal association areas",
                   "Striatum ventral region",
                   "Caudoputamen",
                   "Thalamus",
                   "Medial preoptic nucleus",
                   "Hemispheric regions")
pruneAnatTree(right_tree_for_homology, nodes = nodes_to_merge, method = "BelowNode")

right_tree_leaves <- right_tree_for_homology$Get("name", filterFun = isLeaf) %>% as.data.frame()
```

## Aggregate for the midline
```{r}
# prune nodes of interest:
midline_tree_for_homology <- Clone(midline_tree_clean)
nodes_to_merge<- c("Midbrain", "Medulla", "Pons")
pruneAnatTree(midline_tree_for_homology, nodes = nodes_to_merge, method = "BelowNode")
midline_tree_leaves <- midline_tree_for_homology$Get("name", filterFun = isLeaf) %>% as.data.frame()
```

## Convert mouse volumes to tibble and filter for the left, right, and midline trees
### LEFT
```{r}
mouse_vols_left <- left_tree_for_homology$Get("volumes") %>% as_tibble
filtered_mouse_vols_left <- dplyr::select(mouse_vols_left,"Piriform cortex", 
                   "Subiculum",
                   "Entorhinal area",
                   "Field CA1",
                   "Field CA3",
                   "Dentate gyrus, molecular layer",
                   "Hippocampal region",
                   "Medial amygdalar nucleus",
                   "Cortical subplate",
                   "Anterior cingulate area",
                   "Retrosplenial area",
                   "Primary auditory area",
                   "Agranular insular area",
                   "Perirhinal area",
                   "Primary motor area",
                   "Orbital area, ventrolateral part",
                   "Primary somatosensory area",
                   "Primary visual area",
                   "Temporal association areas",
                   "Posterior parietal association areas",
                   "Pallidum",
                   "left Bed nuclei of the stria terminalis",
                   "Striatum ventral region",
                   "Caudoputamen",
                   "Thalamus",
                   "Medial preoptic nucleus",
                   "Hypothalamus", 
                   "Medial preoptic nucleus",
                   "Hypothalamus",
                   "Hemispheric regions")

mouse_homologs_left<-dplyr::rename(filtered_mouse_vols_left,
                          `Amygdala` = `Cortical subplate`,
                          `Agranular insula`=`Agranular insular area`,
                          `Ventral orbital area` = `Orbital area, ventrolateral part`,
                          `Nucleus accumbens` = `Striatum ventral region`,
                          `Hippocampus`=`Hippocampal region`,
                          `Nucleus accumbens`=`Striatum ventral region`,
                          `Globus pallidus`=`Pallidum`,
                          `Bed nuclei of the stria terminalis` = `left Bed nuclei of the stria terminalis`,
                          `Cerebellar cortex`=`Hemispheric regions`)
        

colnames(mouse_homologs_left)<-sub("^","Left ",colnames(mouse_homologs_left))
```

### RIGHT
## Convert mouse volumes to tibble and filter for the left, right, and midline trees
```{r}
mouse_vols_right <- right_tree_for_homology$Get("volumes") %>% as_tibble
filtered_mouse_vols_right <- dplyr::select(mouse_vols_right, "Piriform cortex", 
                   "Subiculum",
                   "Entorhinal area",
                   "Field CA1",
                   "Field CA3",
                   "Dentate gyrus, molecular layer",
                   "Hippocampal region",
                   "Medial amygdalar nucleus",
                   "Cortical subplate",
                   "Anterior cingulate area",
                   "Retrosplenial area",
                   "Primary auditory area",
                   "Agranular insular area",
                   "Perirhinal area",
                   "Primary motor area",
                   "Orbital area, ventrolateral part",
                   "Primary somatosensory area",
                   "Primary visual area",
                   "Temporal association areas",
                   "Posterior parietal association areas",
                   "Pallidum",
                   "right Bed nuclei of the stria terminalis",
                   "Striatum ventral region",
                   "Caudoputamen",
                   "Thalamus",
                   "Medial preoptic nucleus",
                   "Hypothalamus", 
                   "Medial preoptic nucleus",
                   "Hypothalamus",
                   "Hemispheric regions")

mouse_homologs_right<-dplyr::rename(filtered_mouse_vols_right,
                          `Amygdala` = `Cortical subplate`,
                          `Agranular insula`=`Agranular insular area`,
                          `Ventral orbital area` = `Orbital area, ventrolateral part`,
                          `Nucleus accumbens` = `Striatum ventral region`,
                          `Hippocampus`=`Hippocampal region`,
                          `Nucleus accumbens`=`Striatum ventral region`,
                          `Globus pallidus`=`Pallidum`,
                          `Bed nuclei of the stria terminalis` = `right Bed nuclei of the stria terminalis`,
                          `Cerebellar cortex`=`Hemispheric regions`)

colnames(mouse_homologs_right)<-sub("^","Right ",colnames(mouse_homologs_right))
```

### MIDLINE
## Convert mouse volumes to tibble and filter for the left, right, and midline trees
```{r}
mouse_vols_midline <- midline_tree_for_homology$Get("volumes") %>% as_tibble
filtered_mouse_vols_midline <- dplyr::select(mouse_vols_midline,
                                          "Midbrain", "Medulla", "Pons", "Hindbrain")

filtered_mouse_vols_midline$`Brain stem` <- filtered_mouse_vols_midline$Midbrain + filtered_mouse_vols_midline$Hindbrain 
mouse_homologs_midline <-filtered_mouse_vols_midline

#mouse_homologs_midline <- dplyr::select(filtered_mouse_vols_midline,-"Midbrain", -"Hindbrain")
```

## Calculate total tissue volume
```{r}
vols <- treeVols$Get("volumes") %>% as_tibble
vols_leafs <- treeVols$Get("volumes", filterFun= isLeaf) %>% as_tibble

dat <- vols %>%  cbind(gf)
gf$ventricle_volume <- vols$`ventricular systems`
gf$total_tissue_volume<- dat$root2 - gf$ventricle_volume
gf1 <- dplyr::select(gf, -"ventricle_volume")
```

## Merge and clean up data frames
```{r}
df_filtered_mouse_vols <- gf1 %>%  cbind(mouse_homologs_left) %>%  cbind(mouse_homologs_right) %>%  cbind(mouse_homologs_midline)
```
## Filter mouse data
```{r}
nLimit <- 4
balanced_groups_by_cohort5_1 <-  df_filtered_mouse_vols %>% filter(Is_Wildtype == "WT" & !duplicated(Mouse_ID))
balanced_groups_by_cohort5 <- balanced_groups_by_cohort5_1 %>%  
  dplyr::group_by(Study_Name) %>% 
  dplyr::summarize(nF=sum(Mouse_Sex == "F"),
            nM=sum(Mouse_Sex == "M"),
             n = n()) %>%
  filter (nF > nLimit & nM > nLimit)
balanced_groups_by_cohort5

balanced_groups_by_cohort5 <- balanced_groups_by_cohort5_1 %>% 
  filter (Study_Name %in% balanced_groups_by_cohort5$Study_Name & Treatment_Code %in% c("none", "MOCK", "VEH", "Control","Saline") &
                                                     Background %in% c("C57BL-6J", "C57BL-6N", "C57BL6-N","C57BL-6"))


df_balanced_groups_by_cohort51 <- balanced_groups_by_cohort5 %>% 
          mutate(Mouse_Age = str_remove(Mouse_Age,"P")) %>%
          separate(Mouse_Age, c("a", "b"), fill = "right") %>%
          mutate(a = as.numeric(a), b = as.numeric(b)) %>%
          mutate(Mouse_Age = ifelse(is.na(b) == FALSE, (a + b)/2, 
                                    a), .after = Mouse_Sex) %>%
          dplyr::select(-c(a,b)) 
#Replace NAs with age of 60
df_balanced_groups_by_cohort51["Mouse_Age"][is.na(df_balanced_groups_by_cohort51["Mouse_Age"])] <- 60
df_balanced_groups_by_cohort5 <- df_balanced_groups_by_cohort51%>% filter(Mouse_Age < 200)

df_balanced_groups_by_cohort5$AGE_cent <- df_balanced_groups_by_cohort5$Mouse_Age - mean(df_balanced_groups_by_cohort5$Mouse_Age)

table(df_balanced_groups_by_cohort5$Mouse_Sex, df_balanced_groups_by_cohort5$Study_Name)
table(df_balanced_groups_by_cohort5$Mouse_Sex, df_balanced_groups_by_cohort5$Background)
table(df_balanced_groups_by_cohort5$Background, df_balanced_groups_by_cohort5$Study_Name)
```

# Apply ComBat
```{r}
# Apply ComBat to each background strain
df_just_C57BL6J <-df_balanced_groups_by_cohort5 %>% filter(Background == "C57BL-6J")
table(df_just_C57BL6J$Study_Name, df_just_C57BL6J$Mouse_Sex)

df_just_C57BL6N <-df_balanced_groups_by_cohort5 %>% filter(Background %in% c("C57BL-6N","C57BL6-N"))
table(df_just_C57BL6N$Study_Name, df_just_C57BL6N$Mouse_Sex)
df_just_C57BL6N$Background <- gsub("C57BL6-N", "C57BL-6N", df_just_C57BL6N$Background)

# C57BL6J
mouse_vols_C57BL6J_t <- t(as.matrix(df_just_C57BL6J[,34:95]))
mouse_demo_C57BL6J <- df_just_C57BL6J[,c(1:9,96)]

cohort_batch_C57BL6J <- df_just_C57BL6J$Study_Name 
combat_vols_cohort_C57BL6J <- sva::ComBat(dat = mouse_vols_C57BL6J_t, batch= cohort_batch_C57BL6J)

combat_vols_C57BL6J_reoriented <- t(combat_vols_cohort_C57BL6J)
df_combat_C57BL6J_vols <- mouse_demo_C57BL6J %>%  cbind(combat_vols_C57BL6J_reoriented) %>% as.data.frame()

# C57BL6N
mouse_vols_C57BL6N_t <- t(as.matrix(df_just_C57BL6N[,34:95]))
mouse_demo_C57BL6N <- df_just_C57BL6N[,c(1:9,96)]

cohort_batch_C57BL6N <- df_just_C57BL6N$Study_Name %>% as.factor()
combat_vols_cohort_C57BL6N <- sva::ComBat(dat = mouse_vols_C57BL6N_t, batch = cohort_batch_C57BL6N)

combat_vols_C57BL6N_reoriented <- t(combat_vols_cohort_C57BL6N)
df_combat_C57BL6N_vols <- mouse_demo_C57BL6N %>%  cbind(combat_vols_C57BL6N_reoriented) %>% as.data.frame()

df_combat_vols <- rbind(df_combat_C57BL6J_vols, df_combat_C57BL6N_vols)

saveRDS(df_combat_vols, "/Users/gumae2/Documents/Cross_species_sex_differences/mouse_data/Mouse_homologs_aggregated_postcombat.RDS")
write.csv(df_combat_vols, "/Users/gumae2/Documents/Cross_species_sex_differences/mouse_data/Mouse_homologs_aggregated_postcombat.csv")
```

# Calculate z-scores
```{r}
mouse_vols <- df_combat_vols[,11:ncol(df_combat_vols)] %>% as.matrix()
mouse_demo <- df_combat_vols[,c(1:10)]

zscores_volumes <- matrix(0, nrow = nrow(mouse_vols), ncol = ncol(mouse_vols))
for(j in 1:ncol(mouse_vols)){
  mu <- mean(mouse_vols[,j]) #Then you can compute the ROI-wise averages
  sigma <- sd(mouse_vols[,j]) #And standard deviations
  zscores_volumes[,j] <- (mouse_vols[,j] - mu)/sigma #Then you can use those to make the z-score volumes
}
colnames(zscores_volumes)<-colnames(mouse_vols)
df_zscored_data_homologs <- mouse_demo %>%  cbind(zscores_volumes)
write.csv(df_combat_vols, "/Users/gumae2/Documents/Cross_species_sex_differences/mouse_data/Mouse_homologs_aggregated_postcombat_zscored.csv")
```

### With TBV covariate
Tun the linear model testing for group, correcting for total tissue volume to extract standardized beta values
```{r}
sexdiff_mouse_lm_wTBV <- anatLm(~ Mouse_Sex+Background + total_tissue_volume + AGE_cent , df_zscored_data_homologs,
                             df_zscored_data_homologs[,10:ncol(df_zscored_data_homologs)])
FDR_sexdiff_mouse_wTBV <- anatFDR(sexdiff_mouse_lm_wTBV)
FDR_sexdiff_mouse_wTBV

Sex_mouse_model_outputs_wTBV <- sexdiff_mouse_lm_wTBV %>% as.data.frame()
Sex_mouse_model_outputs_wTBV$label <- rownames(Sex_mouse_model_outputs_wTBV)
write.csv(Sex_mouse_model_outputs_wTBV, file="/Users/gumae2/Documents/Cross_species_sex_differences/cross-species/homologous_ROI_full_mouse_lm_outputs_combat_c57.csv")
```

### Without TBV covariate
Tun the linear model testing for group, correcting for total tissue volume to extract standardized beta values
```{r}
sexdiff_mouse_lm_noTBV <- anatLm(~ Mouse_Sex + AGE_cent + Background, df_zscored_data_homologs,
                             df_zscored_data_homologs[,10:ncol(df_zscored_data_homologs)])
FDR_sexdiff_mouse_noTBV <- anatFDR(sexdiff_mouse_lm_noTBV)
FDR_sexdiff_mouse_noTBV

Sex_mouse_model_outputs_noTBV <- sexdiff_mouse_lm_noTBV %>% as.data.frame()
Sex_mouse_model_outputs_noTBV$label <- rownames(Sex_mouse_model_outputs_noTBV)
write.csv(Sex_mouse_model_outputs_noTBV, file="/Users/gumae2/Documents/Cross_species_sex_differences/cross-species/homologous_ROI_full_mouse_lm_noTTV_combat_c57.csv")
```


# Plots with TBV 
Plot the effects for sex across both species
```{r}
# Merge the human and mouse beta values
Human_sex_model_outputs_wTBV_filtered <- dplyr::select(Human_sex_model_outputs_wTBV, "label", "beta-SexM")
Sex_mouse_model_outputs_wTBV_filtered <- dplyr::select(Sex_mouse_model_outputs_wTBV, "label", "beta-Mouse_SexM")

human_and_mouse_results_wTBV<-merge(Human_sex_model_outputs_wTBV_filtered, Sex_mouse_model_outputs_wTBV_filtered, by="label")
human_and_mouse_results_wTBV$label <- as.character(human_and_mouse_results_wTBV$label)
```

### Correlate effects
Calculate correlations
```{r}
pbcor(human_and_mouse_results_wTBV$`beta-SexM`,human_and_mouse_results_wTBV$`beta-Mouse_SexM`)
```

### Plot results
```{r}
Sex_diffs_plot_wTBV <- ggplot(human_and_mouse_results_wTBV, aes(x=`beta-SexM`, y=`beta-Mouse_SexM`, 
                                                              label=sexdiff_label)) +
  geom_point()+ 
  geom_label_repel(aes(label = sexdiff_label), size = 3, hjust = 0.5, box.padding = 0.5,
                  point.padding=0.5, nudge_x = 0, nudge_y=0, max.overlaps = 45)+
  geom_smooth(method = MASS::rlm) + 
  theme_classic()+xlab("Effect Size Human")+ ylab("Effect Size Mouse")+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14), 
        plot.title = element_text(hjust = 0.5, size = 15, face="bold"),
        legend.position="none")+
 xlim(-0.8,0.8) + ylim(-0.8,0.8)+ 
  geom_vline(xintercept=c(0), linetype="dashed", alpha=0.4)+
  geom_hline(yintercept=c(0), linetype="dashed", alpha=0.4)
Sex_diffs_plot_wTBV
```

## Assign cortical or non cortical label
```{r}
human_and_mouse_results_wTBV$cort_vs_subcort <- human_and_mouse_results_wTBV$label
human_and_mouse_results_wTBV$cort_vs_subcort[1] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[2] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[3] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[4] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[5:8] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[9] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[10:17]<- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[18:25] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[26] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[27] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[28] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[29] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[30:32] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[33] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[34] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[35] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[36:39] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[40] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[41:48] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[49:56] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[57] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[58] <- "Cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[59] <- "Not cortex"
human_and_mouse_results_wTBV$cort_vs_subcort[60] <- "Cortex"
```

### Correlate effects in cortex and subcortex
Calculate correlations
```{r}
cortex_only  <- human_and_mouse_results_wTBV %>% filter(cort_vs_subcort %in% c("Cortex")) 
subcortex_only  <- human_and_mouse_results_wTBV %>% filter(cort_vs_subcort %in% c("Not cortex")) 

pbcor(cortex_only$`beta-SexM`, cortex_only$`beta-Mouse_SexM`)
pbcor(subcortex_only$`beta-SexM`, subcortex_only$`beta-Mouse_SexM`)
```

## PLot results based on cortex/not cortex
```{r}
Sex_diffs_plot_wTBV_cort_subcort <- ggplot(human_and_mouse_results_wTBV, aes(x=`beta-SexM`, y=`beta-Mouse_SexM`, 
                                                              label=label, group=cort_vs_subcort, col=cort_vs_subcort)) +
  geom_point()+ 
  geom_label_repel(aes(label = label), size = 3, hjust = 0.5, box.padding = 0.5,
                  point.padding=0.5, nudge_x = 0, nudge_y=0, max.overlaps = 25)+
  geom_smooth(method = MASS::rlm) +
  theme_classic()+xlab("Effect Size Human")+ ylab("Effect Size Mouse")+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14), 
        plot.title = element_text(hjust = 0.5, size = 15, face="bold"),
        legend.position="bottom")+
 xlim(-0.8,0.8) + ylim(-0.8,0.8)+ 
  scale_fill_manual(values=c("springgreen4", "mediumorchid4"))+ 
  scale_color_manual(values = c("springgreen4", "mediumorchid4"))+
  geom_vline(xintercept=c(0), linetype="dashed", alpha=0.4)+
  geom_hline(yintercept=c(0), linetype="dashed", alpha=0.4)
Sex_diffs_plot_wTBV_cort_subcort
```

