---
title: "HCP_volume_analysis_notwinpairs"
author: "Elisa Guma"
date: "10/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description

In this script I aim to compute maps of sex differences in the brain using Freesurfer outputs from the HCP data.

# Load in necessary libraries
```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RMINC)
library(ggplot2)
library(ggseg)
library(ggsegGlasser)
library(raincloudplots)
library(cowplot)
library(readr)
library(PupillometryR)
library(colortools)
```

## Import data
```{r}
df_HCP_volumes <- readRDS("../input_data/human_anatomy/df_HCP_volumes_clean.RDS")
```

## remove the subject with low euler number
```{r}
df_HCP_volumes <- subset(df_HCP_volumes, df_HCP_volumes$euler > -200)
```

# Remove twin pairs
```{r}
table(df_HCP_volumes$Sex, df_HCP_volumes$ZygositySR)
mz_twins  <- df_HCP_volumes %>% filter(ZygositySR %in% c("MZ"))
table(mz_twins$Sex, mz_twins$Mother_ID)

mz_twin_1 <- read.csv("../input_data/human_anatomy/demographics/mz_twins_1.csv")
df_mz_twin_filtered_1  <- mz_twin_1 %>% filter(QC_include %in% c("1"))
df_mz_twin_filtered <- df_mz_twin_filtered_1[,2] %>% as.data.frame()
colnames(df_mz_twin_filtered) <- "SUBJECT_ID"
df_HCP_volumes_filtered_twins <- merge(df_mz_twin_filtered, df_HCP_volumes, by="SUBJECT_ID")

df_nontwins  <- df_HCP_volumes %>% filter(ZygositySR %in% c("NotMZ", "NotTwin"))

df_HCP_volumes_notwinpairs <- rbind(df_HCP_volumes_filtered_twins, df_nontwins)
```

```{r}
table(df_HCP_volumes_notwinpairs$Sex, df_HCP_volumes_notwinpairs$Race)
table(df_HCP_volumes_notwinpairs$Sex, df_HCP_volumes_notwinpairs$Ethnicity)
table(df_HCP_volumes_notwinpairs$ZygositySR, df_HCP_volumes_notwinpairs$Sex)
```

## Analysis
```{r}
summary(aov(Age_in_Yrs ~ Sex, df_HCP_volumes_notwinpairs))
summary(aov(BMI ~ Sex, df_HCP_volumes_notwinpairs))
summary(aov(euler ~ Sex, df_HCP_volumes_notwinpairs))
summary(aov(SSAGA_Educ ~ Sex, df_HCP_volumes_notwinpairs))
```


# HUMAN Z-SCORE CALCULATION
## define data
```{r}
human_vols <- df_HCP_volumes_notwinpairs[,30:ncol(df_HCP_volumes_notwinpairs)] #Then subset the volumes for the control observations #vols from 54:461
HCP_demo <- df_HCP_volumes_notwinpairs[,1:29]
```

## calculate z-scores
```{r}
zscores_HCP_notwins <- matrix(0, nrow = nrow(human_vols), ncol = ncol(human_vols))
for(j in 1:ncol(human_vols)){
  mu <- mean(human_vols[,j]) #Then you can compute the ROI-wise averages
  sigma <- sd(human_vols[,j]) #And standard deviations
  zscores_HCP_notwins[,j] <- (human_vols[,j] - mu)/sigma #Then you can use those to make the z-score volumes
}
```

## Assign column names and "rebind" demographics data
```{r}
colnames(zscores_HCP_notwins)<-colnames(human_vols)
df_zscores_HCP_notwins <- HCP_demo %>% cbind(zscores_HCP_notwins) 
```


# LINEAR MODEL Z-SCORE 
## Look at total brain, total tissue, and lateral ventricle effects

## Total tissue volume linear model
```{r}
linearmodel <- lm(BrainSegVolNotVent.y ~ Sex + AGE_cent + euler, data=df_zscores_HCP_notwins)
summary(linearmodel)
```

### Plot total tissue volume
```{r}
## z-score total tissue volume
mu_ttv<- mean(df_HCP_volumes_notwinpairs$BrainSegVolNotVent.y)
sigma_ttv <- sd(df_HCP_volumes_notwinpairs$BrainSegVolNotVent.y)
df_HCP_volumes_notwinpairs$zscored_total_tissue_volume <- (df_HCP_volumes_notwinpairs$BrainSegVolNotVent.y - mu_ttv)/sigma_ttv 

ggplot(df_HCP_volumes_notwinpairs, aes(x = Sex, y = zscored_total_tissue_volume, group=Sex, colour=Sex, fill=Sex)) + 
geom_flat_violin(position = position_nudge(x = .1, y = 0),alpha =.4,adjust =0.8)+
  geom_boxplot( width = .1, outlier.shape = NA, alpha = .5) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(side = "l", range_scale = .4, alpha = .3) +
  #Adjust theme
  scale_fill_manual(values=c("F"="deepskyblue","M"="goldenrod1"))+
    scale_colour_manual(values=c("F"="deepskyblue","M"="goldenrod1"))+
    theme_classic()+
  labs(
    y="Total Tissue Volume (Z-scored)",
    fill="Sex")+  theme(text = element_text(size = 14))+ylim(-4, 4)+
  scale_y_continuous(
    sec.axis = sec_axis(name=expression(Total ~ Tissue ~ Volume~ (mm^3)), ~ (. * sigma_ttv) + mu_ttv), 
    name="Total Tissue Volume (Z-scored)", limits=c(-4,4)) +
  coord_cartesian(xlim = c(1.2, NA), clip = "off")
```

# Run across all regions

### Remove white matter and ventricle volume from data fram
```{r}
df_zscores_HCP_notwins_greymatter <-dplyr::select(df_zscores_HCP_notwins, -"total_white_matter",-"total_ventricle_vol",-starts_with("CC_"),
                                          -ends_with("Ventricle"), -ends_with("Inf.Lat.Vent"),
                                          -ends_with("CortexVol"), -ends_with("WhiteMatterVol"), 
                                          -ends_with("SubCortGrayVol"), -"BrainSegVol", -"Left.Hippocampus", -"Right.Hippocampus",
                                          -"Left.Amygdala", -"Right.Amygdala",-ends_with("Cerebellum.White.Matter"),-"CSF", 
                                          -"Brain.Stem", -"eTIV", -"TotalGrayVol", -ends_with("VentralDC"))
```

# Run across all ROIs with TBV as a covariate
### Full ROI subset
```{r}
Sex_diffs_allROI_wTBV_zscore= anatLm(~ Sex + AGE_cent + BrainSegVolNotVent.y, df_zscores_HCP_notwins_greymatter,df_zscores_HCP_notwins_greymatter[,30:ncol(df_zscores_HCP_notwins_greymatter)])
FDR_Sex_diffs_allROI_wTBV_zscore <- anatFDR(Sex_diffs_allROI_wTBV_zscore)
FDR_Sex_diffs_allROI_wTBV_zscore

# calculate t-values
tvals <- abs(Sex_diffs_allROI_wTBV_zscore[, 'tvalue-SexM'])
dof <- 959
pvals_sex<- 2*pt(q = tvals, df = dof, lower.tail = FALSE) %>% as_tibble()

tvals <- abs(Sex_diffs_allROI_wTBV_zscore[, 'tvalue-AGE_cent'])
dof <- 959
pvalsAGE_cent <- 2*pt(q = tvals, df = dof, lower.tail = FALSE) %>% as_tibble()

tvals <- abs(Sex_diffs_allROI_wTBV_zscore[, 'tvalue-BrainSegVolNotVent.y'])
dof <- 959
pvalsBrainSegVolNotVent.y <- 2*pt(q = tvals, df = dof, lower.tail = FALSE) %>% as_tibble()

# aggregate model outputs 
model_outputs_Sex_diffs_wTBV <- Sex_diffs_allROI_wTBV_zscore %>% as.data.frame()
model_outputs_Sex_diffs_wTBV$label <- rownames(Sex_diffs_allROI_wTBV_zscore) 
model_outputs_Sex_diffs_wTBV1 <- cbind(model_outputs_Sex_diffs_wTBV, pvals_sex,pvalsAGE_cent,pvalsBrainSegVolNotVent.y,FDR_Sex_diffs_allROI_wTBV_zscore)
colnames(model_outputs_Sex_diffs_wTBV1) <- c("F-statistic","R-squared","beta-(Intercept)","beta-SexM", "beta-AGE_cent","beta-BrainSegVolNotVent.y", "tvalue-(Intercept)","tvalue-SexM", "tvalue-AGE_cent", "tvalue-BrainSegVolNotVent.y","logLik","label","pvalsdx_groupXYY_H","pvalsAGE_cent","pvalsBrainSegVolNotVent.y", "qvalue-F-statistic", "qvalue-tvalue-(Intercept)", "qvalue-tvalue-SexM", "qvalue-tvalue-AGE_cent","qvalue-tvalue-BrainSegVolNotVent.y")
```
## Turn lm results into a df (with TBV correction) 
Here we can threshold the t-values to those above 5%FDR correction (and use those for the Beta values as well). This is for the TBV corrected findings
```{r}
model_outputs_Sex_diffs_wTBV$label <- gsub("_ROI_volume", "", model_outputs_Sex_diffs_wTBV$label)
model_outputs_Sex_diffs_wTBV$label <- gsub("Right.", "Right-", model_outputs_Sex_diffs_wTBV$label)
model_outputs_Sex_diffs_wTBV$label <- gsub("Left.", "Left-", model_outputs_Sex_diffs_wTBV$label)
#model_outputs_Sex_diffs_wTBV$label <- gsub("Right-Cerebellum.Cortex", "right-cerebellum-white-matter",  model_outputs_Sex_diffs_wTBV$label)
#model_outputs_Sex_diffs_wTBV$label <- gsub("Left-Cerebellum.Cortex", "left-cerebellum-white-matter",  model_outputs_Sex_diffs_wTBV$label)
model_outputs_Sex_diffs_wTBV$label <- gsub("Right-Cerebellum.Cortex", "right-cerebellum-cortex", model_outputs_Sex_diffs_wTBV$label)
model_outputs_Sex_diffs_wTBV$label <- gsub("Left-Cerebellum.Cortex", "left-cerebellum-cortex", model_outputs_Sex_diffs_wTBV$label)
model_outputs_Sex_diffs_wTBV$label <- gsub("Whole_brainstem", "brain-stem", model_outputs_Sex_diffs_wTBV$label)
model_outputs_Sex_diffs_wTBV$label <- gsub("Whole_hippocampus", "Hippocampus", model_outputs_Sex_diffs_wTBV$label)
model_outputs_Sex_diffs_wTBV$label <- gsub("Whole_amygdala", "Amygdala", model_outputs_Sex_diffs_wTBV$label)

## Threshold at 5% FDR
sig_tvalues_pos_Sex_wTBV <- model_outputs_Sex_diffs_wTBV$`tvalue-SexM`> 2.39
significant_positive_tvalues_sex_wTBV <-model_outputs_Sex_diffs_wTBV[sig_tvalues_pos_Sex_wTBV,]
sig_tvalues_neg_Sex_wTBV <- model_outputs_Sex_diffs_wTBV$`tvalue-SexM`< -2.39
significant_negative_tvalues_sex_wTBV <-model_outputs_Sex_diffs_wTBV[sig_tvalues_neg_Sex_wTBV,]
sig_tvalues_Sex_wTBV <- model_outputs_Sex_diffs_wTBV[sig_tvalues_pos_Sex_wTBV | sig_tvalues_neg_Sex_wTBV, ]
```

## Plot cortical effects with TBV correction
```{r}
#significant
ggplot(sig_tvalues_Sex_wTBV) +   
  geom_brain(atlas = glasser, 
             colour="white", position = position_brain(hemi +side ~ .),
             mapping=aes(fill=`beta-SexM`)) + theme_void() +
   theme(legend.text = element_text(size = 10))+
    labs(title = "Sex diffs (with TTV correction)", fill="Beta")+
 scale_fill_gradient2(low = c("navyblue","blue","royalblue","dodgerblue1","deepskyblue1","turquoise3"),
                        mid = NULL,
                        high = c("goldenrod1","orange1","firebrick2", "firebrick3","firebrick","red4"),
                        midpoint = 0,na.value="grey",limits = c(-1.5,1.5), 
                        n.breaks=5)
#unthresholded
ggplot(model_outputs_Sex_diffs_wTBV) +   
  geom_brain(atlas = glasser, 
             colour="white", position = position_brain(hemi +side ~ .),
             mapping=aes(fill=`beta-SexM`)) + theme_void() +
   theme(legend.text = element_text(size = 10))+
    labs(title = "Sex diffs (with TTV correction)", fill="T-value")+
 scale_fill_gradient2(low = c("navyblue","blue","royalblue","dodgerblue1","deepskyblue1","turquoise3"),
                        mid = NULL,
                        high = c("goldenrod1","orange1","firebrick2", "firebrick3","firebrick","red4"),
                        midpoint = 0,na.value="grey",limits = c(-1.5,1.5), 
                        n.breaks=5)
```

## Subcortical
### Axial
```{r}
#significant
ggplot(model_outputs_Sex_diffs_wTBV) +   
  geom_brain(atlas = aseg, side="coronal",
               mapping=aes(fill=`beta-SexM`), show.legend=T) +
  theme_void() + 
  theme(legend.text = element_text(size = 10))+
    labs(title = "Sex diffs (with TTV correction)", fill="Beta")+
 scale_fill_gradient2(low = c("navyblue","blue","royalblue","dodgerblue1","deepskyblue1","turquoise3"),
                        mid = NULL,
                        high = c("goldenrod1","orange1","firebrick2", "firebrick3","firebrick","red4"),
                        midpoint = 0,na.value="grey",limits = c(-1.5,1.5), 
                        n.breaks=5)
#unthresholded
ggplot(sig_tvalues_Sex_wTBV) +   
geom_brain(atlas = aseg, side="coronal",
               mapping=aes(fill=`beta-SexM`), show.legend=T) +
  theme_void() + 
    theme(legend.text = element_text(size = 10))+
    labs(title = "Sex diffs (with TTV correction)", fill="T-value")+
 scale_fill_gradient2(low = c("navyblue","blue","royalblue","dodgerblue1","deepskyblue1","turquoise3"),
                        mid = NULL,
                        high = c("goldenrod1","orange1","firebrick2", "firebrick3","firebrick","red4"),
                        midpoint = 0,na.value="grey",limits = c(-6,6), 
                        n.breaks=5)
```

### Sagittal
```{r}
#significant
ggplot(sig_tvalues_Sex_wTBV) +   
  geom_brain(atlas = aseg, side="sagittal",
               mapping=aes(fill=`beta-SexM`), show.legend=T) +
  theme_void() + 
  theme(legend.text = element_text(size = 10))+
    labs(title = "Sex diffs (with TTV correction)", fill="Beta")+
 scale_fill_gradient2(low = c("navyblue","blue","royalblue","dodgerblue1","deepskyblue1","turquoise3"),
                        mid = NULL,
                        high = c("goldenrod1","orange1","firebrick2", "firebrick3","firebrick","red4"),
                        midpoint = 0,na.value="grey",limits = c(-2,2), 
                        n.breaks=5)
```

## Correlate effect sizes from Lmer and regular lm
```{r}
df_human_results_wTBV <- readRDS("../input_data/human_anatomy/df_human_results_wTBV_regularLm.RDS")

df_human_results_notwins_wTBV <- model_outputs_Sex_diffs_wTBV %>% dplyr::select(c(`label`, `beta-SexM`))
colnames(df_human_results_notwins_wTBV) <- c("label", "beta-sex-notwins")
df_human_results_comparison <- merge(df_human_results_wTBV, df_human_results_notwins_wTBV, by="label")

pbcor(df_human_results_comparison$`beta-sex-regular-lm`, df_human_results_comparison$`beta-sex-notwins`)

ggplot(df_human_results_comparison, aes(x=`beta-sex-regular-lm`, y=`beta-sex-notwins`)) +
  geom_point()+  geom_smooth(method=lm)  + 
  theme_classic()+xlab("Effect Size Lm")+ ylab("Effect Size No Twins")+
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14), 
        plot.title = element_text(hjust = 0.5, size = 15, face="bold"),
        legend.position="bottom")+
 xlim(-0.8,0.8) + ylim(-0.8,0.8)+ 
  scale_fill_manual(values=c("springgreen4", "mediumorchid4"))+ 
  scale_color_manual(values = c("springgreen4", "mediumorchid4"))+
  geom_vline(xintercept=c(0), linetype="dashed", alpha=0.4)+
  geom_hline(yintercept=c(0), linetype="dashed", alpha=0.4)
```

# Run liner mixed-effects model across all ROIs

```{r}
df_zscores_HCP_notwins_greymatter$Father_ID <- as.factor(df_zscores_HCP_notwins_greymatter$Father_ID)
df_zscores_HCP_notwins_greymatter$Mother_ID <- as.factor(df_zscores_HCP_notwins_greymatter$Mother_ID)

Sex_diffs_allROI_wTBV_zscore_lmer= anatLmer(~ Sex + AGE_cent + BrainSegVolNotVent.y +euler + (1|Father_ID), df_zscores_HCP_notwins_greymatter,df_zscores_HCP_notwins_greymatter[,30:ncol(df_zscores_HCP_notwins_greymatter)])
Sex_diffs_allROI_wTBV_zscore_DF <- anatLmerEstimateDF(Sex_diffs_allROI_wTBV_zscore_lmer)
FDR_Sex_diffs_allROI_wTBV_zscore_lmer <- anatFDR(Sex_diffs_allROI_wTBV_zscore_DF)
FDR_Sex_diffs_allROI_wTBV_zscore_lmer

lmer_outputs_Sex_diffs_wTBV <- Sex_diffs_allROI_wTBV_zscore_lmer %>% as.data.frame()
lmer_outputs_Sex_diffs_wTBV$label <- rownames(Sex_diffs_allROI_wTBV_zscore_lmer) 
```
### Turn lmer results into a df (with TBV correction) 
Here we can threshold the t-values to those above 5%FDR correction (and use those for the Beta values as well). This is for the TBV corrected findings
```{r}
lmer_outputs_Sex_diffs_wTBV$label <- gsub("_ROI_volume", "", lmer_outputs_Sex_diffs_wTBV$label)
lmer_outputs_Sex_diffs_wTBV$label <- gsub("Right.", "Right-", lmer_outputs_Sex_diffs_wTBV$label)
lmer_outputs_Sex_diffs_wTBV$label <- gsub("Left.", "Left-", lmer_outputs_Sex_diffs_wTBV$label)
lmer_outputs_Sex_diffs_wTBV$label <- gsub("Right-Cerebellum.Cortex", "right-cerebellum-white-matter",  lmer_outputs_Sex_diffs_wTBV$label)
# lmer_outputs_Sex_diffs_wTBV$label <- gsub("Left-Cerebellum.Cortex", "left-cerebellum-white-matter",  lmer_outputs_Sex_diffs_wTBV$label)
# lmer_outputs_Sex_diffs_wTBV$label <- gsub("Right-Cerebellum.Cortex", "right-cerebellum-cortex", lmer_outputs_Sex_diffs_wTBV$label)
lmer_outputs_Sex_diffs_wTBV$label <- gsub("Left-Cerebellum.Cortex", "left-cerebellum-cortex", lmer_outputs_Sex_diffs_wTBV$label)
lmer_outputs_Sex_diffs_wTBV$label <- gsub("Whole_brainstem", "brain-stem", lmer_outputs_Sex_diffs_wTBV$label)
lmer_outputs_Sex_diffs_wTBV$label <- gsub("Whole_hippocampus", "Hippocampus", lmer_outputs_Sex_diffs_wTBV$label)
lmer_outputs_Sex_diffs_wTBV$label <- gsub("Whole_amygdala", "Amygdala", lmer_outputs_Sex_diffs_wTBV$label)

## Threshold at 5% FDR
sig_tvalues_pos_Sex_wTBV_lmer <- lmer_outputs_Sex_diffs_wTBV$`tvalue-SexM`> 2.47
significant_positive_tvalues_sex_wTBV_lmer <-lmer_outputs_Sex_diffs_wTBV[sig_tvalues_pos_Sex_wTBV_lmer,]
sig_tvalues_neg_Sex_wTBV_lmer <- lmer_outputs_Sex_diffs_wTBV$`tvalue-SexM`< -2.47
significant_negative_tvalues_sex_wTBV_lmer <-lmer_outputs_Sex_diffs_wTBV[sig_tvalues_neg_Sex_wTBV_lmer,]
sig_tvalues_Sex_wTBV_lmer <- lmer_outputs_Sex_diffs_wTBV[sig_tvalues_pos_Sex_wTBV_lmer | sig_tvalues_neg_Sex_wTBV_lmer, ]
```

## Plotting brain results
```{r}
#beta
ggplot(sig_tvalues_Sex_wTBV_lmer) +   
  geom_brain(atlas = glasser, 
             colour="white", position = position_brain(hemi +side ~ .),
             mapping=aes(fill=`beta-SexM`)) + theme_void() +
   theme(legend.text = element_text(size = 10))+
    labs(title = "Sex diffs (with TTV correction)", fill="Beta")+
 scale_fill_gradient2(low = c("navyblue","blue","royalblue","dodgerblue1","deepskyblue1","turquoise3"),
                        mid = NULL,
                        high = c("goldenrod1","orange1","firebrick2", "firebrick3","firebrick","red4"),
                        midpoint = 0,na.value="grey",limits = c(-1.5,1.5), 
                        n.breaks=5)
#unthresholded
ggplot(lmer_outputs_Sex_diffs_wTBV) +   
  geom_brain(atlas = glasser, 
             colour="white", position = position_brain(hemi +side ~ .),
             mapping=aes(fill=`beta-SexM`)) + theme_void() +
   theme(legend.text = element_text(size = 10))+
    labs(title = "Sex diffs (with TTV correction)", fill="T-value")+
 scale_fill_gradient2(low = c("navyblue","blue","royalblue","dodgerblue1","deepskyblue1","turquoise3"),
                        mid = NULL,
                        high = c("goldenrod1","orange1","firebrick2", "firebrick3","firebrick","red4"),
                        midpoint = 0,na.value="grey",limits = c(-1.5,1.5), 
                        n.breaks=5)

#tvalue
ggplot(sig_tvalues_Sex_wTBV_lmer) +   
  geom_brain(atlas = glasser, 
             colour="white", position = position_brain(hemi +side ~ .),
             mapping=aes(fill=`tvalue-SexM`)) + theme_void() +
   theme(legend.text = element_text(size = 10))+
    labs(title = "Sex diffs (with TTV correction)", fill="T-value")+
 scale_fill_gradient2(low = c("navyblue","blue","royalblue","dodgerblue1","deepskyblue1","turquoise3"),
                        mid = NULL,
                        high = c("goldenrod1","orange1","firebrick2", "firebrick3","firebrick","red4"),
                        midpoint = 0,na.value="grey",limits = c(-6,6), 
                        n.breaks=5)


```

## Subcortical results with TBV correction

### Axial
```{r}
#Beta
ggplot(sig_tvalues_Sex_wTBV_lmer) +   
  geom_brain(atlas = aseg, side="axial",
               mapping=aes(fill=`beta-SexM`), show.legend=T) +
  theme_void() + 
  theme(legend.text = element_text(size = 10))+
    labs(title = "Sex diffs (with TTV correction)", fill="Beta")+
 scale_fill_gradient2(low = c("navyblue","blue","royalblue","dodgerblue1","deepskyblue1","turquoise3"),
                        mid = NULL,
                        high = c("goldenrod1","orange1","firebrick2", "firebrick3","firebrick","red4"),
                        midpoint = 0,na.value="grey",limits = c(-1.5,1.5), 
                        n.breaks=5)
```

### Sagittal
```{r}
ggplot(sig_tvalues_Sex_wTBV_lmer) +   
  geom_brain(atlas = aseg, side="sagittal",
               mapping=aes(fill=`beta-SexM`), show.legend=T) +
  theme_void() + 
  theme(legend.text = element_text(size = 10))+
    labs(title = "Sex diffs (with TTV correction)", fill="Beta")+
 scale_fill_gradient2(low = c("navyblue","blue","royalblue","dodgerblue1","deepskyblue1","turquoise3"),
                        mid = NULL,
                        high = c("goldenrod1","orange1","firebrick2", "firebrick3","firebrick","red4"),
                        midpoint = 0,na.value="grey",limits = c(-1.5,1.5), 
                        n.breaks=5)

```
