---
title: "Mouse_expression_data_cleanup"
author: "Elisa Guma"
date: "5/17/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load in necessary librarie

library(tidyverse)
library(RMINC)
library(magrittr)
library(broom)
library(data.tree)
library(ggrepel)
library(ggplot2)
library(ggseg)
library(cowplot)
library(readr)
library(sva)
source('../input_data/tree_tools.R')
```

#Description
This script is used to aggregate gene expression data from the Allen Mouse Brain Atlas
into the homologous brain regions we define in this study, based on the DSURQE atlas

#Load data
```{r}
mouse_expression_matrix <- as_tibble(data.table::fread("../input_data/raw_data_files/mouse_expression_matrix.csv", header = TRUE))
```
# Selection regions
```{r}
df_mouse_expression_matrix <- dplyr::filter(mouse_expression_matrix, Region %in% c("left Piriform cortex", "right Piriform cortex", 
                                                                                   "left Subiculum","right Subiculum",
                                                                                   "left Entorhinal area","right Entorhinal area",
                                                                                   "left Field CA1","right Field CA1",
                                                                                   "left Field CA3", "right Field CA3",
                                                                                   "left Field CA2, stratum radiatum", "right Field CA2, stratum radiatum",
                                                                                   "left Field CA2, stratum oriens", "right Field CA2, stratum oriens",
                                                                                   "left Field CA2, pyramidal layer", "right Field CA2, pyramidal layer",
                                                                                   "left Dentate gyrus, molecular layer","right Dentate gyrus, molecular layer",
                                                                                   "left Hippocampus","right Hippocampus",
                                                                                   "left Medial amygdalar nucleus", "right Medial amygdalar nucleus",
                                                                                   "left Cortical subplate","right Cortical subplate",
                                                                                   "left Anterior cingulate area","right Anterior cingulate area",
                                                                                   "left Retrosplenial area","right Retrosplenial area",
                                                                                   "left Primary auditory area","right Primary auditory area",
                                                                                   "left Agranular insular area", "right Agranular insular area",
                                                                                   "left Perirhinal area","right Perirhinal area",
                                                                                   "left Primary motor area","right Primary motor area",
                                                                                   "left Orbital area, ventrolateral part","right Orbital area, ventrolateral part",
                                                                                   "left Primary somatosensory area","right Primary somatosensory area",
                                                                                   "left Primary visual area","right Primary visual area",
                                                                                   "left Temporal association areas","right Temporal association areas",
                                                                                   "left Posterior parietal association areas","right Posterior parietal association areas",
                                                                                   "left Pallidum, dorsal region","right Pallidum, dorsal region",
                                                                                   "left Pallidum, caudal region","right Pallidum, caudal region",
                                                                                   "left Pallidum, ventral region","right Pallidum, ventral region",
                                                                                   "left Pallidum, medial region","right Pallidum, medial region",
                                                                                   "left Bed nuclei of the stria terminalis","right Bed nuclei of the stria terminalis",
                                                                                   "left Striatum ventral region","right Striatum ventral region",
                                                                                   "left Caudoputamen","right Caudoputamen",
                                                                                   "left Thalamus","right Thalamus",
                                                                                   "left Medial preoptic nucleus","right Medial preoptic nucleus",
                                                                                   "left Cerebellar cortex", "right Cerebellar cortex",
                                                                                   "Midbrain", "Medulla", "Pons"))

df_mouse_expression_matrix$Region <- gsub("left Field CA1", "left Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("right Field CA1", "right Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("left Field CA3", "left Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("right Field CA3", "right Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("left Field CA2, stratum radiatum", "left Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("right Field CA2, stratum radiatum", "right Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("left Field CA2, stratum oriens", "left Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("right Field CA2, stratum oriens", "right Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("left Field CA2, pyramidal layer", "left Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("right Field CA2, pyramidal layer", "right Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("left Dentate gyrus, molecular layer", "left Hippocampus", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("right Dentate gyrus, molecular layer", "right Hippocampus", df_mouse_expression_matrix$Region)

df_mouse_expression_matrix$Region <- gsub("Pallidum, dorsal region", "Pallidum", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("Pallidum, caudal region", "Pallidum", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("Pallidum, ventral region", "Pallidum", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("Pallidum, medial region", "Pallidum", df_mouse_expression_matrix$Region)
df_mouse_expression_matrix$Region <- gsub("Bed nuclei of the stria terminalis", "Pallidum", df_mouse_expression_matrix$Region)

df_mouse_expression_matrix_agg <- df_mouse_expression_matrix %>% filter(!is.na(Region)) %>% group_by(Region) %>% summarise_all(mean)
```

## Aggregate hypothalamus separately
```{r}
df_mouse_expression_matrix_hypothal <- dplyr::filter(mouse_expression_matrix, Region %in% c("left Hypothalamus-other", "right Hypothalamus-other"))
df_mouse_expression_matrix_hypothal_agg <- df_mouse_expression_matrix_hypothal %>% filter(!is.na(Region)) %>% group_by(Region) %>% summarise_all(mean)
df_mouse_expression_matrix_hypothal_agg$Region <- gsub("Hypothalamus-other", "Hypothalamus", df_mouse_expression_matrix_hypothal_agg$Region)
```

## Aggregate BNST separately
```{r}
df_mouse_expression_matrix_bnst <- dplyr::filter(mouse_expression_matrix, Region %in% c("left Bed nuclei of the stria terminalis", "right Bed nuclei of the stria terminalis"))
df_mouse_expression_matrix_bnst_agg <- df_mouse_expression_matrix_bnst %>% filter(!is.na(Region)) %>% group_by(Region) %>% summarise_all(mean)
#df_mouse_expression_matrix_bnst_agg$Region <- gsub("Hypothalamus-other", "Hypothalamus", df_mouse_expression_matrix_hypothal_agg$Region)
```

## Aggregate hippocampus separately
```{r}
df_mouse_expression_matrix_hipp_subfields <- dplyr::filter(mouse_expression_matrix, Region %in% c("left Field CA1", "right Field CA1",
                                                           "left Field CA3", "right Field CA3","left Dentate gyrus, molecular layer","right Dentate gyrus, molecular layer"))

df_mouse_expression_matrix_hipp_subfields_agg <- df_mouse_expression_matrix_hipp_subfields %>% filter(!is.na(Region)) %>% group_by(Region) %>% summarise_all(mean)
```

## Rename regions
```{r}
df_mouse_expression_matrix_agg$Region <- gsub("Cortical subplate", "Amygdala", df_mouse_expression_matrix_agg$Region)
df_mouse_expression_matrix_agg$Region <- gsub("Cortical subplate", "Amygdala", df_mouse_expression_matrix_agg$Region)
df_mouse_expression_matrix_agg$Region <- gsub("Agranular insular area", "Agranular insula", df_mouse_expression_matrix_agg$Region)
df_mouse_expression_matrix_agg$Region <- gsub("Orbital area, ventrolateral part", "Ventral orbital area", df_mouse_expression_matrix_agg$Region)
df_mouse_expression_matrix_agg$Region <- gsub("Striatum ventral region", "Nucleus accumbens", df_mouse_expression_matrix_agg$Region)
df_mouse_expression_matrix_agg$Region <- gsub("Pallidum", "Globus pallidus", df_mouse_expression_matrix_agg$Region)
df_mouse_expression_matrix_agg$Region <- gsub("Striatum ventral region", "Nucleus accumbens", df_mouse_expression_matrix_agg$Region)
```

## Aggregate brainstem ROIs
```{r}
brainstem_ROIS <- df_mouse_expression_matrix_agg[1:3,2:ncol(df_mouse_expression_matrix_agg)] %>% as_tibble()
row_sum <- colMeans(brainstem_ROIS[,1:ncol(brainstem_ROIS)]) 

df_mouse_expression_matrix_aggregate <- rbind(df_mouse_expression_matrix_agg[,2:ncol(df_mouse_expression_matrix_agg)],row_sum)
Region <- c(df_mouse_expression_matrix_agg$Region,"Brainstem")
df_mouse_expression_matrix_aggregate1 <- cbind(Region,df_mouse_expression_matrix_aggregate) %>% as_tibble()
```

## Add hypothalamus and subfieldROIs
```{r}
df_mouse_expression_matrix_aggregate_hp_1 <- rbind(df_mouse_expression_matrix_aggregate1, df_mouse_expression_matrix_hypothal_agg)
df_mouse_expression_matrix_aggregate_hp <- rbind(df_mouse_expression_matrix_aggregate_hp_1, df_mouse_expression_matrix_bnst_agg)
df_mouse_expression_matrix_aggregate_final <- rbind(df_mouse_expression_matrix_aggregate_hp, df_mouse_expression_matrix_hipp_subfields_agg)
```

# Save data
These outputs are available in the github folder
```{r}
#write.csv(df_mouse_expression_matrix_aggregate_final, "../input/gene_expression/aggregated_data/aggregated_mouse_gene_expression.csv")
#saveRDS(df_mouse_expression_matrix_aggregate_final, "../input/gene_expression/aggregated_data/aggregated_mouse_gene_expression.RDS")
```

#filter by homologous genes
```{r}
homologous_genes <- read.csv("../input/gene_expression/raw_data_files/MouseHumanGeneHomologs.csv")
homologous_genes$Mouse

homologs <- homologous_genes$Mouse %>% as.character() 

justgenes <- df_mouse_expression_matrix_aggregate_final[,2:ncol(df_mouse_expression_matrix_aggregate_final)]
t_justgenes <- t(justgenes)
rownames(t_justgenes) <- colnames(justgenes)
colnames(t_justgenes) <- df_mouse_expression_matrix_aggregate_final$Region
gene <- rownames(t_justgenes) %>% as.character()
t_justgenes1 <- cbind (gene, t_justgenes) %>% as_tibble()

t_genes_homologs<- subset(t_justgenes1, gene %in% homologs)

#write.csv(t_genes_homologs, "../input/gene_expression/aggregated_data/transposed_homologous_mouse_genes.csv")
#saveRDS(t_genes_homologs, "../input/gene_expression/aggregated_data/transposed_homologous_mouse_genes.RDS")
```

